% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[Ligatures=TeX]{Cambria}
  \setmonofont[]{Consolas}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{xurl}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Frontier and Localhost: How Production AI Learns Outside the Weights},
  pdfauthor={Mikhail L. Arbuzov; Alexey A. Shvets; Sisong Bei},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Frontier and Localhost: How Production AI Learns Outside the
Weights}
\author{Mikhail L. Arbuzov \and Alexey A. Shvets \and Sisong Bei}
\date{}

\begin{document}
\maketitle

\textbf{Abstract}

Reliability has moved from the weights to the scaffold. Surveying
production and research LLM systems from 2023--2026, we show that
practitioners repeatedly wrap frozen frontier models in local,
persistent, feedback-updated artifacts: instructions, skills, memories,
tools, orchestration graphs, and governance pipelines. We argue that
these artifacts form a deployment-time learning layer --- a
\emph{localhost scaffold} --- that adapts a general model to the
recurring failure topology of a specific patch. We formalise this as
\textbf{artifact-layer descent}: failures provide noisy loss signals,
candidate scaffold deltas are generated, gates accept or reject them,
accepted deltas persist and may be promoted across contexts. The
resulting two-loop architecture explains the convergence of modern agent
systems across IDE plugins, vertical-bundle vendors, and self-evolving
research agents, and it reveals an unfilled design corner: automated
local scaffold evolution combined with governed, auditable,
cross-organisation promotion --- a \emph{DP-FedAvg-style aggregator with
an automated eval gate at the scaffold layer}. The contribution is
conceptual and survey-based. A practitioner-facing companion paper will
handle reference architectures, maturity ladders, design principles, and
the full failure taxonomy.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. The gradient moved outside the
model}\label{the-gradient-moved-outside-the-model}

The standard mental model of an LLM system is the model. Pre-training
fixes the weights; fine-tuning or RLHF nudges them; deployment wraps a
thin prompt-plus-tool layer around the result and ships. Errors are
weight errors; reliability is a weight-training problem.

Production has stopped behaving this way. In 2025--2026, frontier models
update on cycles measured in months, but the systems built on top of
them update continuously --- in markdown files, persistent memories,
retrieved skills, registered tools, evaluation suites, agent topologies,
version-controlled prompt repositories, and PR-gated governance. The
fast learning is no longer in the weights. It is in a fast-changing
\emph{localhost} scaffold wrapped around a slow-changing frontier.

This paper argues that \emph{the localhost is doing learning}. Not
merely metaphorically: scaffold artifacts are updated from failure
signal, the updated artifacts reduce future failure probability on a
local patch, and the update process --- observe failure, generate
candidate delta, gate it, persist it, optionally promote it --- has the
structure of stochastic descent. The substrate happens to be discrete,
symbolic, and human-readable rather than continuous and differentiable,
but the function is the same.

Once you accept the framing, the field snaps into place. A Claude Skill
is not ``just a markdown file.'' It is a learned procedural weight kept
outside the model. A Cursor Rule is not ``just an instruction.'' It is a
local parameter for one project's distribution shift. A RAG connector is
not ``just retrieval.'' It is patch-specific memory access. A tool
registry is not ``just integration.'' It is capability provisioning
along an axis the base model cannot reach. A PR review against a shared
rule repository is not ``just governance.'' It is a promotion gate on a
federated update. An eval suite that blocks merge is not ``just
testing.'' It is a validation-loss check before promotion. A
cross-tenant artifact repository is not ``just sharing.'' It is
federated learning across distributed shards. The mechanisms were
already there; only the name was missing.

We survey 142 systems across six substrates (instructions, skills,
memory, tools, orchestration, governance), score each on a 0--5
patch-plasticity rubric, and analyse the fourteen that satisfy a strict
\emph{composite two-loop} criterion: an inner loop that updates a
persisted artifact from session feedback, an outer loop that promotes
the artifact across scopes under explicit versioning. The two-loop
design space is a 3×3 matrix on gate type --- \emph{Auto-gated} when the
commit-or-promote decision is operationalised by an algorithmic
criterion, \emph{Human-gated} when it is reviewer judgment. The matrix
has four cells of interest, three populated by working systems and one
empty by engineering logic. The fourth diagonal --- fully automated,
governed, cross-tenant --- has no surveyed occupant. That empty corner
is the paper's open-question headline and a concrete Paper-4 target.

The architectural review of how to actually build such systems --- five
reference patterns, a five-level maturity model, a substrate-selection
rubric, three migration paths, the full nine-category failure taxonomy,
and eight design principles --- is deferred to follow-on practitioner
work. This paper is the survey and the framework.

\subsection{2. Why patch errors must live outside the
weights}\label{why-patch-errors-must-live-outside-the-weights}

The scaffold is not merely where production systems happen to store
patches. It is the \emph{only reasonable substrate} for patch-specific
adaptation, because the weight substrate is structurally wrong for it.
Frontier training and production reliability optimise for objectives
that are in tension, and that tension forces adaptation outside the
model.

\textbf{Frontier training optimises for breadth.} The weights are tuned
for cross-patch generalisation: they smooth over local conventions,
contradictory tenant preferences, and rare workflow-specific failures in
order to preserve broad competence across millions of unseen
deployments. The smoothing is not incidental. It is the training
objective. A model that overfits to one tenant's naming conventions or
one team's preferred error-handling style loses transfer to the next
tenant; a model that encodes every patch's idiosyncratic eval
expectation loses the general-purpose competence that made it worth
deploying in the first place.

\textbf{Production reliability is achieved in depth.} Inside a single
deployment patch, reliability is determined by exactly the stuff global
training had to smooth away: local naming conventions, team-specific
workflows, domain-specific failure patterns, project-specific schemas,
idiosyncratic tool chains, customer-specific risk tolerances, recurring
local mistakes, institutional preferences, hidden evaluator
expectations. The per-patch quantities from our previous work ---
failure-mode catalogue size \(|C_D|\), hard-fraction \(\beta_D\),
mode-discovery rate \(\sigma_D\) --- are local by construction, and a
library calibrated to one patch under-covers the next.

\textbf{Encoding every patch in the weights is structurally infeasible.}
Three constraints rule out the obvious approach. \emph{Economic:}
millions of patches multiplied by billions of parameters times the cost
of per-patch fine-tuning is not a viable training-compute budget.
\emph{Release-cycle:} the patch wants daily updates; weight releases
ship quarterly at best. \emph{Cross-patch interference:} overfitting the
weights to one tenant's preferences degrades behaviour for adjacent
tenants whose preferences contradict. The weights are tuned for
cross-patch invariants by construction; that is what they are
\emph{for}.

The localhost scaffold solves the conflict. It is the only substrate
that can:

\begin{itemize}
\tightlist
\item
  overfit aggressively to one tenant without damaging another;
\item
  encode local conventions without polluting global behaviour;
\item
  update daily instead of waiting for release cycles;
\item
  be inspected, versioned, audited, and rolled back;
\item
  attach tools, memories, rules, workflows, and evals that are
  meaningless outside the patch;
\item
  specialise inside the patch while the model remains broadly competent.
\end{itemize}

Each scaffold substrate targets a different patch-residual class.
Instructions and tools address \emph{capability} residuals --- the model
could not do this on its own (PAL for arithmetic, RAG for retrieval,
constrained decoders for format). Skills and memory address
\emph{catalogue} residuals --- the same patch-specific failure mode
keeps recurring and now has a patch. Orchestration addresses
\emph{hard-fraction} residuals --- the per-step decision was too hard,
so specialised sub-agents bring it back into capability range.
Governance addresses \emph{propagation} residuals --- a correction
proven in one deployment is wasted unless it can be promoted to others
under audit.

Our previous work (Arbuzov, Shvets \& Bei, 2025; \emph{Architecture of
Errors}, 2026) argued that LLM errors concentrate sparsely at key tokens
and cluster into a finite catalogue of recurring modes whose size grows
polylogarithmically in observed failures. The framework left a
\emph{placement} question --- where do the interventions live? The
answer follows from the tension above. The interventions live in
scaffold artifacts because they cannot live anywhere else. The frontier
model generalises; the localhost scaffold specialises; reliability comes
from governing that specialisation.

\subsection{3. Artifact-layer descent}\label{artifact-layer-descent}

The remaining question is whether ``scaffold evolution'' is learning in
any technical sense or just engineering practice. The answer depends on
what one means by learning.

Let \(D\) be a deployment patch --- a specific application, user, or
task distribution. Let \(M\) be the frozen frontier model, accessed via
inference but not updated. Let \(S_t\) be the scaffold at time \(t\):
the layered collection \(\{S_t^{(1)}, \ldots, S_t^{(6)}\}\) of
instruction set, skill library, memory store, tool registry,
orchestration graph, and governance configuration. Let
\(\pi(M, S_t, x)\) denote the system's behaviour on input \(x\) --- the
model running on the model-plus-scaffold composite. Let
\(L_D(S_t) = \mathbb{E}_{x \sim D}[\ell(\pi(M, S_t, x))]\) be the
expected residual failure rate on the patch.

A failure observation produces a candidate delta \(\Delta S\): an
instruction edit, a new skill, a memory correction, a tool
re-registration, a topology change, or a governance-rule update. A gate
\(G\) accepts \(\Delta S\) if estimated patch loss decreases or if risk
stays within tolerance:

\[
S_{t+1} = S_t + G(\Delta S)
\]

Different surveyed systems instantiate \(G\) differently. NanoResearch's
gate is a semantic-merge by the orchestrator (algorithmic). SkillRL's
gate is an RL threshold at task-category accuracy below 0.4
(algorithmic). AlphaEvolve's gate is a cascade evaluator with MAP-Elites
archive selection (algorithmic). SkillForge's gate is an LLM-judge with
\textgreater90\% human agreement \emph{plus} a human support engineer
(hybrid algorithmic + reviewer). OpenCore's gate is a reviewer-judgment
check against \texttt{core-update-gate.md}, with no algorithmic
threshold --- only a checklist (purely reviewer-judgment). DSPy's gate
is gradient-like over prompt-template space without persistent
commitment. Reflexion's gate is essentially absent: any reflection
becomes a memory entry.

Different estimators of the same quantity. Each surveyed system is
asking, by some method, whether \(\Delta S\) moves \(S_t\) toward lower
\(L_D\). The substrate is discrete: scaffold edits are typically
textual, procedural, or symbolic, not differentiable in the
backpropagation sense. The gate is mediated: humans, evals, LLM-judges,
or surrogate verifiers stand in for the gradient. The update is
governed: not every candidate \(\Delta S\) propagates; some are
rejected, some are quarantined to a single user's fork, some climb a
layered hierarchy from local to shared scope before they take.

We call this \emph{artifact-layer descent}. In the limit where scaffold
edits are represented as coordinates in an artifact basis and the gate
is a differentiable validation-loss estimator, the process is gradient
descent. In ordinary deployments, where edits are discrete and gates are
eval-, judge-, or reviewer-mediated, it is a stochastic approximation to
descent. The two are continuous with one another; the difference is one
of substrate and signal-to-noise, not of mechanism.

Two loops follow naturally. \textbf{Loop 1} is within-context: a session
feedback signal triggers an update to a persisted local artifact.
\textbf{Loop 2} is cross-context: a local artifact is promoted to a
shared scope (cross-project, cross-user, cross-organisation) under
explicit versioning. Loop 1 is the local SGD step; Loop 2 is federated
averaging across distributed shards. Each loop has a gate; gates are
\textbf{Auto-gated} when the decision is an operationalised algorithmic
criterion (RL threshold, eval score, LLM-judge with quantified
agreement, surrogate verifier, benchmark cascade), \textbf{Human-gated}
when the decision is reviewer judgment (automation often precedes the
gate but the gate itself is reviewer discretion), and \textbf{None} when
the loop is absent.

These five terms --- Loop 1, Loop 2, Auto-gated, Human-gated, None ---
anchor the rest of the paper.

The framing positions the work against several adjacent literatures.
CoALA (Sumers et al., 2023) proposes a static cognitive architecture of
modular memory and structured action; we update CoALA's snapshot into a
dynamic deployment architecture with governance. The self-evolution
surveys (Tao 2024; Gao et al.~2026; Fang et al.~2025) treat agents as
learners that update weights, representations, or optimiser-selected
artifacts --- they locate the gradient inside the model. The
memory-survey of Du (2026) and the skills-survey of Zhou et al.~(2026)
cover individual substrates with version-and-promotion machinery that
mirrors S3 and S2 here but stops at the substrate boundary. ACE (2025)
treats context as a single evolving playbook. Each adjacent work touches
a piece of the picture; none develops the cross-substrate artifact-layer
descent framing or the design space it implies.

\subsubsection{3.1 The mechanism stack}\label{the-mechanism-stack}

The abstract update rule \(S_{t+1} = S_t + G(\Delta S)\) expands into
six operations that every patch-plastic system must perform --- whether
explicitly engineered or accidentally emergent.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Residual capture.} Log failures, corrections, rejected
  outputs, user edits, eval regressions, tool-call failures. The Loop-1
  signal source. Without it, \(\Delta S\) has no provenance.
\item
  \textbf{Mode discovery.} Cluster repeated failures into patch-specific
  error modes. The Paper-2 catalogue, instantiated locally per patch.
\item
  \textbf{Delta synthesis.} Convert a recurring mode into a candidate
  scaffold edit: an instruction rewrite, a new skill, a memory
  correction, a tool binding, an orchestration change, an eval case.
  This is the production of \(\Delta S\).
\item
  \textbf{Delta validation.} Estimate whether the edit reduces \(L_D\)
  without collateral damage on adjacent inputs. This is the gate \(G\)
  --- Auto-gated when the check is an algorithmic criterion (RL
  threshold, eval score, LLM-judge with quantified agreement),
  Human-gated when it is reviewer judgment.
\item
  \textbf{Promotion control.} Decide whether the accepted \(\Delta S\)
  stays in the local fork or climbs the scope hierarchy --- project,
  stack, organisation, cross-tenant. This is Loop 2.
\item
  \textbf{Drift and bloat control.} Expire stale rules, prune
  conflicting memories, roll back regressed skills, detect over-fitting
  to one patch. Without this layer, accumulated \(\Delta S\) produces
  the scaffold-level analogue of plasticity loss.
\end{enumerate}

Each operation is the locus of a separate engineering discipline that
does not yet exist in coordinated form. Surveyed systems implement the
operations partially and unevenly: research full-auto systems excel at
(3) and (4), production governance leaders at (5), industry hybrids at
most of (1)--(5) but rarely (6). Implementation patterns for each of the
six are deferred to follow-on work; this paper's purpose is to make the
discipline visible.

\subsubsection{3.2 The mechanism made formal: scaffold updates as
discrete stochastic
descent}\label{the-mechanism-made-formal-scaffold-updates-as-discrete-stochastic-descent}

The mechanism stack above makes artifact-layer descent operational, but
the word ``descent'' still needs to earn its keep. A skeptical reading
is available: scaffold updates are just engineering edits, not
gradients; PR review is not backpropagation; an LLM-judge is not a
derivative; a markdown rule is not a differentiable parameter. All true.
But they do not defeat the claim. They locate the claim in the correct
optimization class.

Scaffold evolution is not backpropagation through continuous weights. It
is discrete, gated, patch-local stochastic descent over a structured
artifact space.

Let the frontier model have fixed parameters \(\theta_M\). Let the
scaffold be a structured parameter object

\[\theta_s \in \Theta_s = \Theta_1 \times \Theta_2 \times \Theta_3 \times \Theta_4 \times \Theta_5 \times \Theta_6,\]

where the six coordinates correspond to the six substrates of §4:
instructions, skills, memory, tools, orchestration, and governance. A
coordinate may be a line in a \texttt{CLAUDE.md} file, a Cursor rule, a
skill file, a memory entry, a retrieval configuration, a tool schema, an
agent-routing edge, an eval threshold, or a promotion rule. The space
\(\Theta_s\) is not smooth. It is discrete, symbolic, and versioned. But
it is still a parameter space: changing \(\theta_s\) changes the
behavior of the model--scaffold composite.

For a deployment patch \(D\), define patch loss as

\[L_D(\theta_s) = \mathbb{E}_{x \sim D}\left[\ell(\pi(\theta_M, \theta_s, x))\right],\]

where \(\pi(\theta_M,\theta_s,x)\) is the output or action trace
produced by the frozen frontier model under scaffold \(\theta_s\), and
\(\ell\) is an operational loss. Depending on the patch, \(\ell\) may be
a hard-failure indicator, user-correction rate, eval-suite error,
tool-call failure, retrieval miss, schema violation, reviewer
disagreement, or graded task score. The important distinction is
substrate: \(\theta_M\) is fixed during deployment; \(\theta_s\) is the
object being fit.

A failure observation does not directly yield an analytic gradient. It
yields a candidate edit. Let

\[z_t \in \mathcal{A}(\theta_s^t)\]

be an admissible scaffold edit proposed at time \(t\), drawn from a
proposal process

\[z_t \sim Q_t(z \mid H_t),\]

where \(H_t\) is the accumulated history of failures, corrections,
rejected outputs, eval regressions, reviewer comments, tool-call traces,
and prior accepted deltas. The admissible edit set
\(\mathcal{A}(\theta_s^t)\) is the local neighborhood of the current
scaffold: the rule can be rewritten, a skill added, a memory corrected,
a tool schema tightened, an orchestration edge changed, an eval case
inserted, a promotion policy modified.

Because the space is discrete, the relevant gradient analogue is not
\(\nabla_{\theta_s} L_D\). It is the finite directional loss change
induced by an edit:

\[\Delta_z L_D(\theta_s) = L_D(\theta_s \oplus z) - L_D(\theta_s),\]

where \(\theta_s \oplus z\) denotes the scaffold obtained by applying
edit \(z\). A useful edit is one for which

\[\Delta_z L_D(\theta_s) < 0.\]

The gate estimates whether this inequality holds. Let
\(\widehat{\Delta}_z L_D\) be the gate's noisy estimate of the
directional loss change. In an eval-gated system,
\(\widehat{\Delta}_z L_D\) is measured on a held-out task suite. In an
LLM-judge system, it is estimated by model-mediated scoring, ideally
calibrated against human agreement. In a reviewer-gated system, it is
estimated by human credit assignment and PR review. In an RL-threshold
system, it is estimated by reward improvement. In a verifier-gated
system, it is estimated by a symbolic, learned, or surrogate verifier.

The update rule is therefore

\[\theta_s^{t+1} = \begin{cases} \theta_s^t \oplus z_t, & \text{if } G_t(z_t)=1, \\ \theta_s^t, & \text{if } G_t(z_t)=0, \end{cases}\]

where the gate \(G_t\) accepts when the estimated improvement clears the
deployment's threshold:

\[G_t(z_t)=1 \quad \Longleftrightarrow \quad \widehat{\Delta}_{z_t} L_D(\theta_s^t) + \lambda\,\Omega(z_t) \leq -\tau_t.\]

Here \(\Omega(z_t)\) is an optional complexity or risk penalty ---
longer instructions, broader tool permissions, larger memory writes,
more invasive topology rewrites, wider promotion scope --- while
\(\lambda\) controls regularization and \(\tau_t\) is the required
improvement margin. Production systems often implement this without
writing the equation: an eval suite blocks a merge if regression
appears; a reviewer rejects a vague rule; a governance system prevents a
local artifact from becoming shared until evidence accumulates.

The implicit learning rate is edit scope. A narrow one-line rule patch
is a small step. A full rubric rewrite is a large step. A memory
correction local to one user is a small step. Promotion from USER to
PROJECT to STACK to CORE is an increase in update radius. Cross-tenant
promotion is the largest step and therefore requires the strongest gate.
This is why governance is not administrative overhead. It is the
learning-rate and regularization system for scaffold descent.

The correspondence to standard optimization is therefore not decorative:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2683}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7317}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Optimization concept
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Scaffold analogue
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Parameter vector & Instructions, skills, memories, tools, routing
graphs, eval gates, governance rules \\
Forward pass & Model--scaffold composite executing on patch input \\
Loss & Failure, user correction, eval regression, retrieval miss, tool
error, schema violation \\
Directional derivative & Finite loss change \(\Delta_z L_D(\theta_s)\)
induced by a candidate edit \\
Gradient estimate & Credit assignment to the scaffold coordinate
responsible for the failure \\
Backpropagation analogue & Tracing output failure backward through
prompt, retrieval, memory, tool, or workflow layer \\
SGD step & Accepted scaffold edit \\
Mini-batch & Multiple recurring failures aggregated before update \\
Learning rate & Edit aggressiveness and promotion scope \\
Validation loss & Held-out patch eval before merge \\
Regularization & Complexity penalty, bloat control, cross-patch
regression checks \\
Overfitting & Scaffold improves on this patch and degrades elsewhere \\
Rollback & Reverting a harmful accepted update \\
\end{longtable}
}

The backpropagation row requires care. Scaffold systems do not compute
chain-rule derivatives through markdown files or tool registries. But
they do perform credit assignment through a pipeline. If the final
answer fails, the responsible coordinate may be a retrieval
configuration three layers upstream, a missing tool schema constraint,
an underspecified instruction, a stale memory, or a bad orchestration
edge. Engineers do this manually. LLM-judges and surrogate verifiers do
it semi-automatically. The operation is backpropagation-like in function
--- output error is assigned backward to an upstream parameter --- but
not backpropagation in the differentiable-weight sense.

Several concrete cases make the mechanism visible.

In a RAG system, suppose failures recur because retrieved context omits
the decisive clause. The observed loss is not merely ``the model
hallucinated.'' Credit assignment traces the error to the retrieval
coordinate: chunking, query rewrite, metadata filter, embedding model,
reranker, citation policy, or context-packing rule. The candidate edit
might tighten chunk boundaries, add a metadata constraint, change the
reranker, or insert a regression eval containing the missed clause. The
gate tests whether retrieval recall and answer accuracy improve on the
patch eval without degrading adjacent cases. If accepted, the scaffold
has moved downhill on the local loss surface.

In a pitch-review workflow, suppose the system repeatedly overrates
decks that claim a ``huge market'' without numeric evidence. The failure
cluster identifies an under-specified rubric coordinate. The candidate
edit changes ``evaluate TAM size'' into ``require a numeric market
estimate and source; if absent, cap TAM score at 4/10.'' The gate runs
the revised rubric against held-out decks with human scores. If
agreement improves without damaging unrelated criteria, the edit is
committed. The rubric is not commentary. It is a scaffold parameter
updated in response to loss.

In a tool-using agent, suppose calls repeatedly fail because optional
parameters are underspecified. The responsible coordinate may be the
tool schema, the call policy, or the clarification rule. The candidate
edit adds required fields, constrained decoding, or a pre-call
uncertainty check. The gate estimates whether tool-call failure
decreases without increasing refusal, latency, or unnecessary
clarification. If accepted, the scaffold has fit a recurring local
failure mode.

In a memory system, suppose a stale customer preference is repeatedly
retrieved and causes bad decisions. The failure is not a weight error.
The responsible coordinate is a memory entry plus its provenance,
priority, and expiry policy. The candidate edit corrects, expires, or
version-tags the memory. The gate checks whether the correction improves
future behavior without erasing useful history. Again: loss, credit
assignment, candidate update, validation, commit.

Under mild conditions, this is descent in the formal sense. Assume a
stationary patch \(D\), bounded loss \(0 \leq L_D(\theta_s) \leq B\),
and a gate with positive descent bias:

\[\mathbb{E}\!\left[\Delta_{z_t} L_D(\theta_s^t) \mid G_t(z_t)=1,\theta_s^t\right] \leq -\gamma_t\]

for accepted nontrivial edits, up to finite gate noise. Then

\[\mathbb{E}\!\left[L_D(\theta_s^{t+1}) \mid \theta_s^t\right] \leq L_D(\theta_s^t) - p_t \gamma_t + \varepsilon_t,\]

where \(p_t = \Pr(G_t(z_t)=1)\) and \(\varepsilon_t\) captures false
accepts, finite-eval noise, evaluator bias, and distribution drift. If
harmful accepts are controlled so that
\(\sum_t \varepsilon_t < \infty\), and if useful local edits remain
reachable until no improving edit exists, the process converges to a
local \(G\)-stable scaffold: a scaffold for which the proposal-and-gate
mechanism can no longer find an accepted edit with negative expected
directional loss. This is not a global optimum of all possible
scaffolds. It is the ordinary target of local stochastic optimization:
no reachable accepted step improves the current loss estimate.

In non-stationary patches, \(D = D_t\), the same mechanism tracks a
moving optimum rather than converging once. This is why recency
weighting, replay, pruning, expiry, and rollback are not optional
hygiene. They are the scaffold-level analogues of continual-learning
machinery. Without them, accumulated edits create scaffold bloat, stale
rules, memory pollution, and patch-level plasticity loss.

This formalization also clarifies the role of overfitting. In weight
training, overfitting is usually a defect: the model becomes too
specialized to the training distribution and loses transfer. In scaffold
adaptation, local overfitting is partly the point. Frontier weights are
trained to preserve cross-patch generality; they must smooth over local
conventions, rare workflow-specific failures, contradictory tenant
preferences, and hidden evaluator expectations. A deployment patch needs
the opposite: selective specialization to local residuals. The scaffold
is therefore an intentionally overfittable layer, but one whose
overfitting is scoped, inspectable, versioned, gated, and reversible. A
scaffold that improves one repository, hospital workflow, legal
diligence process, or support desk may degrade performance elsewhere.
That is not a contradiction. It is patch adaptation. The mistake is
promoting that edit beyond its evidence radius.

Loop 1 and Loop 2 are now mathematically interpretable. Loop 1 proposes
and validates local descent steps. Loop 2 controls the radius over which
an accepted local step is allowed to generalize. USER-level acceptance
is a small-radius update. PROJECT-level promotion is wider. STACK-level
promotion is wider still. CORE or cross-tenant promotion is the largest
and riskiest step. The two-loop design space is therefore not just a
taxonomy of engineering patterns. It is a taxonomy of how systems
estimate, accept, regularize, and generalize scaffold descent.

The artifact-layer descent framing is mechanism when five conditions
hold: measurable patch loss, residual capture, failure-to-coordinate
credit assignment, candidate delta synthesis, and a gate that estimates
directional loss change before persistence or promotion. It is merely
metaphor when those conditions are absent. This distinction explains the
survey split. Research full-auto systems automate delta synthesis and
validation but often lack governance. Production governance systems
validate and promote but often lack failure-triggered candidate
generation. Hybrid systems close more of the loop. The empty enterprise
corner is the case where all pieces must hold across tenants.

The conclusion is simple: the frontier model is trained to avoid
overfitting across patches. The scaffold exists to overfit safely inside
one.

\subsection{4. Six scaffold substrates}\label{six-scaffold-substrates}

A patch-plastic scaffold composes six substrates. They are the
coordinates of \(S_t\).

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Substrate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
What persists
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Score-5 mark
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{S1 --- Instructions} & Project/team/org rules attached to a
workspace & Versioned, two-loop, Auto-gated promotion \\
\textbf{S2 --- Skills} & Procedural artifacts (code/text) loaded on
demand & Skill bank self-modifies on failure, unit-test gated,
versioned, promoted \\
\textbf{S3 --- Memory} & Cross-session experience (episodic / semantic /
procedural) & Provenance + versioning + correction pathways \\
\textbf{S4 --- Tools} & Vetted action surface + context schemas &
Domain-curated bundle is the product; eval-gated tool versions \\
\textbf{S5 --- Orchestration} & Multi-role graph + routing &
Population/policy updates from outer-loop signal \\
\textbf{S6 --- Governance} & Versioning, promotion, audit, rollback &
dev→staging→prod with eval thresholds + audit log + rollback \\
\end{longtable}
}

The 0--5 rubric is uniform across substrates: \textbf{0} ephemeral;
\textbf{1} persistent local; \textbf{2} persistent + tool attachment, no
feedback; \textbf{3} multi-scope or feedback but no automated promotion;
\textbf{4} automated feedback updates the scaffold (one loop closed);
\textbf{5} two-loop versioned promotion plus governance (both loops
closed).

Each substrate maps to a different Paper-2 residual class. S1 and S4
carry capability provisioning. S2 and S3 accumulate the patch-specific
failure-mode catalogue. S5 lowers per-step hard-fraction by
specialisation. S6 is the gate substrate --- without it, none of the
other five can be updated safely across deployments. \emph{The
substrates are not a taxonomy of nice-to-haves. They are the layers of
the scaffold parameter vector, separately addressable by Loop 1 and Loop
2.} Score distributions and per-substrate exemplars are summarised in
§6.1; the maturity-rubric details for practitioner adoption are deferred
to follow-on work.

\subsection{5. Survey method and evidence
tiers}\label{survey-method-and-evidence-tiers}

The survey covers 142 LLM systems published or productionised between
January 2023 and May 2026, split into two sets. The \textbf{core corpus}
of 104 candidate scaffold-learning systems satisfies the patch-plastic
discriminator (score ≥ 3 on at least one substrate, indicating skills,
memory, or multi-scope persistence beyond ephemeral prompts). The
\textbf{contrast corpus} of 38 systems is commonly described in the
public discourse as agentic or self-improving but fails at least one
discriminator --- typically Loop 2 absent, no persistent artifact
distinct from optimiser state, or in-context-only adaptation. Headline
statistics quote the core corpus; the contrast set is preserved as
asterisked entries in the matrices for definitional clarity.

\textbf{Inclusion.} Any system that (a) wraps a foundation model with at
least one persistent artifact updated from deployment signal, or (b) is
named in the public discourse as a ``self-improving,'' ``scaffolded,''
or ``agentic'' system. The contrast set (Self-Refine, ReAct, plain Mem0,
DSPy/TextGrad, etc.) is included precisely so the discriminator has
something to mark against.

\textbf{Exclusion.} Systems whose only update mechanism is weight-level
fine-tuning or RLHF --- these are model-side, not scaffold-side, and
belong to the self-evolution literature this paper differentiates from.
Systems with no public evidence of deployment (white papers without
code, demos, or production claim).

\textbf{Evidence tiers.} Each system carries one or more of: (T1)
peer-reviewed publication, (T2) arXiv preprint, (T3) production claim
with metrics, (T4) open-source artifact (GitHub, registry), (T5)
industry-blog or interview-level documentation. Systems with T1 or T4
evidence carry highest weight in the cluster analysis (§7); T3 systems
are admissible where the claim is concrete (specific benchmark, named
integration target); T5 systems are admissible only when corroborated by
T1--T4 elsewhere.

\textbf{Scoring.} Each system received a substrate-by-substrate 0--5
score against the rubric of §4 and an integrated patch-plasticity (PP)
score in \{0,1,2,3,3.5,4,4.5,5\}. Half-scores were used sparingly for
systems where one criterion was clearly met and another partially.
Ambiguous cases were resolved conservatively (downward). Each scoring
decision is recorded in the master CSV.

\textbf{Composite two-loop criterion.} A separate, stricter audit
applied to 22 candidate systems. A system qualifies as composite
two-loop if and only if it satisfies both: Loop 1 modifies a persisted
artifact from a session signal \emph{and} Loop 2 has an explicit
cross-context promotion mechanism with versioning or lineage. Fourteen
systems passed the audit; eight failed (MAE, MetaGen, MetaReflection,
CLIN --- Loop-2 absent; DSPy, TextGrad --- no persistent artifact
distinct from optimiser state; Mem0, Zep, ChatGPT Memory --- no Loop-2
promotion path).

\textbf{Audit trail.} The full per-system spreadsheet
(\texttt{p3\_master\_scores.csv}), per-substrate evidence harvests, and
the composite-two-loop audit
(\texttt{p3\_harvest\_composite\_two\_loop.md}) are in the paper's
repository. Reviewers can re-run the analysis from primary sources.

\subsection{6. Survey findings: the scaffold is unevenly
mature}\label{survey-findings-the-scaffold-is-unevenly-mature}

The 104-system core corpus does not mature uniformly. Tools and
integrated systems are the most mature substrates; skills and
orchestration lag; instructions are ubiquitous but shallow; memory is
the most paradoxical of the six. This section presents the headline
distributions before the cluster interpretation of §7. The full
per-system spreadsheet is in \texttt{p3\_master\_scores.csv}; Appendix A
lists representative rows.

\subsubsection{6.1 Substrate maturity is
uneven}\label{substrate-maturity-is-uneven}

Aggregating the survey by substrate gives this distribution:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1400}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.0400}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.0500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2200}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.5500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Substrate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(n\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Mean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Score-5 systems
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Survey finding
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{S1 --- Instructions} & 12 & 2.83 & \emph{(none)} & Rules persist
across IDEs and agents (Claude Code's CLAUDE.md, Cursor Rules,
AGENTS.md, Windsurf, Continue.dev) but no surveyed instruction system
closes both an automated inner loop and a governed two-loop promotion.
Ceiling held at 4 by Claude Code and Replit \\
\textbf{S2 --- Skills} & 16 & 2.62 & SAGE, COSPLAY & Research frontier
is active (failure-triggered skill update); production governance is
weaker. Anthropic's skills repository is one-pool with PR review but no
automated eval gate \\
\textbf{S3 --- Memory} & 19 & 3.16 & SSGM & Cross-session memory is
widespread; \emph{correctable, versioned} memory is rare. Production
memory systems (ChatGPT Memory, Claude Memory, Mem0, Zep, MemGPT) treat
memory as append-only \\
\textbf{S4 --- Tools} & 19 & 3.63 & MCP, Harvey, Hippocratic AI & Most
mature production substrate. Tool bundles are the commercial unit; MCP
is the cross-vendor standard with 9,400+ public servers and 78\%
enterprise adoption \\
\textbf{S5 --- Orchestration} & 19 & 2.68 & MAE, Memento-Skills &
Multi-agent topologies exist (LangGraph, AutoGen, CrewAI) but topology
\emph{evolution} is rare; production rarely exposes governed promotion
paths for emerging agent graphs \\
\textbf{S6 --- Governance} & 21 & 3.05 & Braintrust, Vellum, LangSmith
Hub, AGENTS.md/AAIF & High production maturity for prompts-as-code ---
immutable commits, eval-gated PR merge, sub-five-minute rollback ---
typically without inner-loop adaptation \\
\textbf{INTEGRATED} & 34 & 3.74 & NanoResearch, AutoAgent, SkillRL & The
high-water mark when present, but the cluster is dominated by research
systems lacking enterprise governance \\
\end{longtable}
}

The unevenness is the survey's central empirical pattern. The field has
independently converged on the six substrates, but no system matures all
six at once. Research systems learn fast and lack enterprise governance;
production systems govern well and learn slowly; open standards solve
artifact portability but not adaptation; vertical-bundle vendors solve
patch specificity but rarely expose cross-tenant promotion. The missing
architecture is not ``agents with tools'' or ``memory with evals.'' It
is \emph{governed scaffold adaptation} --- fast local learning paired
with safe cross-context promotion.

\subsubsection{6.2 Representative systems by
cluster}\label{representative-systems-by-cluster}

Five clusters recur across the corpus. The four-cluster grouping below
maps each system to its gate types and to the structural lesson it
carries. The matrix in §8 places the same systems on the 3×3 design
space.

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1400}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1400}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.4200}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Cluster
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Representative systems (year, PP score)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Loop 1 / Loop 2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
What the cluster proves
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Research full-auto & NanoResearch (2026, 5), SkillRL (2026, 5),
AlphaEvolve (2025, 4), DGM (2025, 4), ADAS (2024, 4), Voyager (2023, 4)
& Auto / Auto & Both loops can be fully automated under algorithmic
gates; benchmark gains compound; enterprise governance is uniformly
absent \\
Production hybrid & SkillForge (2026, 4), CASCADE (2025, 4.5), Sierra OS
2.0 (2025, 3), Cognition Devin (2024, 3) & Auto / Human & Algorithmic
Loop-1 gate plus reviewer-judgment Loop-2 gate is the production-viable
cell; SkillForge is the architectural exemplar \\
Reviewer-judgment & OpenCore (2025, 4) & Human / Human & Automated
extraction machinery with reviewer-judgment gates on both loops; the
only published \(n>1\) mini-batch rule in the corpus (5-dream
threshold); cross-organisation federation via
fork-and-contribute-back \\
Governance-first & Braintrust, Vellum, LangSmith Hub (all 2024, S6 = 5)
& None / Auto & Eval-gated promotion exists; candidate generation is
manual (engineers iterate prompts); no patch-plastic inner loop \\
Open standards & AGENTS.md (2025), MCP (2024), Claude Skills (2025),
Cursor Rules (2024) & Mixed / Mixed & Vendor-cross convergence on
artifact format and tool attachment; 60k+ AGENTS.md repos; 9,400+ MCP
servers; 78\% enterprise MCP adoption \\
\end{longtable}
}

\subsubsection{6.3 Surprising absences}\label{surprising-absences}

What is missing from the survey is as instructive as what is present.
Five absences stand out across the 104-system core corpus:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Failure-triggered memory is rare in production.} Deployed
  memory systems prefer recording preferences and successful facts; only
  Reflexion, the Voyager skill library, Gemini Code Assist's
  PR-rejection memory, and the classical SOAR chunking pattern
  explicitly record residuals. The most obvious Loop-1 signal is the one
  production memory systems mostly ignore.
\item
  \textbf{Cross-customer skill promotion is absent.} AGENTS.md is
  cross-vendor; Anthropic's skills repository is one-pool. No surveyed
  system has a vetted cross-organisational marketplace with eval-gated
  promotion.
\item
  \textbf{Memory versioning is mostly unsolved.} Only SSGM provides
  explicit provenance, versioning, and correction pathways. Production
  default is append-only with implicit retrieval-based forgetting --- at
  odds with the patch-plastic discriminator's correction requirement.
\item
  \textbf{Scaffold-level curriculum is absent.} No surveyed system
  selects \emph{which} sessions inform the gradient. Every session
  contributes equally; weighting by recency, severity, or
  representativeness has not been published.
\item
  \textbf{Overfitting detection at Loop-2 promotion is absent.} No
  surveyed system runs a held-out eval set per project to catch
  scaffolds that over-fit one deployment before they propagate to
  another.
\end{enumerate}

Each absence is a Paper-4 target. The conclusion echoes them; the survey
makes them visible.

\subsection{7. Convergent patterns}\label{convergent-patterns}

The 142-system corpus, scored against the rubric of §4, produces four
archetypal clusters that recur across both research and industry. We
present the clusters here rather than walking every system; the full
table is in the repository.

\textbf{Research full-auto.} Systems where both loops fire from
algorithmic criteria and no human gates the routine update.
\textbf{NanoResearch} (Xu et al., 2026) co-evolves Skills, Memory, and
Policy with a semantic-merge orchestrator; Innovation 4.96 → 5.65 and
Compliance 6.66 → 8.96 across three rounds. \textbf{SkillRL}
(arXiv:2602.08234) uses RL reward distillation with a hard threshold at
task-category accuracy below 0.4 and reports +15.3\% over strong
baselines. \textbf{AlphaEvolve} (DeepMind, 2025) runs a Gemini-driven
ensemble with a cascade evaluator and a MAP-Elites archive, deployed at
Google to gain +0.7\% Borg compute worldwide, +23\% on the Gemini
training kernel, +32.5\% on FlashAttention, and a first improvement over
Strassen's 1969 bound on 4×4 complex matrix multiplication (48 scalar
multiplications). \textbf{DGM} (Zhang et al., 2025) modifies its own
source code with archive-based parent selection; SWE-bench 20\% → 50\%.
The cluster maximises velocity. Enterprise governance is uniformly
absent.

\textbf{Production hybrid.} Systems where Loop 1 is Auto-gated (an
algorithmic criterion) and Loop 2 is Human-gated (reviewer approval
before propagation). \textbf{SkillForge} (Alibaba Cloud,
arXiv:2604.08618) drives auto-generated skill diffs through an LLM-judge
with \textgreater90\% human agreement (the Loop-1 gate); human support
engineers then review what the LLM-judge passes (the Loop-2 gate). VFS
version commits provide full lineage. \textbf{CASCADE} (Huang et al.,
2025) reports the corpus's largest ablation gain --- 93.3\% vs.~35.4\%
on SciSkillBench --- driven by memory consolidation gated by human-agent
collaboration. \textbf{Sierra Agent OS 2.0} routes AI-driven Insights
through human-authored Expert Answers in GitHub-style Workspaces.
\textbf{Cognition Devin} ships dynamic in-session tool creation with a
67\% PR-merge rate (the merged PRs are the Loop-2 promotion). This is
the production-viable cell.

\textbf{Governance-first.} Systems where Loop 2 is Auto-gated by an eval
criterion but Loop 1 is essentially absent --- there is no automatic
candidate-generation pipeline. \textbf{Braintrust}, \textbf{Vellum},
\textbf{LangSmith Hub}, \textbf{W\&B Weave}, \textbf{Agenta},
\textbf{LangFuse}. The pattern is \emph{prompts-as-code}: immutable
commits, semantic-version tag pointers, PR-blocked merges on eval
regression, sub-five-minute rollback. Engineers iterate the artifacts
manually; the eval pipeline gates the promotion. We frame this as
\textbf{MLOps for scaffolds}, distinct from MLOps for weights.

\textbf{Open standards.} Systems where the artifact format and the
layering convention are open standards across vendors.
\textbf{AGENTS.md} (Linux Foundation, December 2025) is now in over
60,000 GitHub repositories; Lulla et al.~(2026) report −28.64\%
wall-clock and −16.58\% tokens with no quality loss after adding an
AGENTS.md scaffold. \textbf{MCP} (Anthropic, 2024) has 9,400+ public
servers, 78\% enterprise adoption, \textasciitilde97M SDK
downloads/month. \textbf{Claude Agent Skills}, \textbf{Cursor Rules},
\textbf{GitHub Copilot custom instructions}, \textbf{Windsurf rules},
\textbf{Continue.dev rules}, \textbf{Zed AI rules}, \textbf{Replit
\texttt{replit.md}} all converge on the same pattern: a git-tracked
markdown artifact attached to a workspace, with an MCP-style tool layer
for actions. Convergence across vendors is the strongest evidence the
architecture is real rather than a single vendor's local optimum.

The four clusters do not partition the corpus --- many systems sit
between two clusters --- but they capture the architectural variation.
Vertical-bundle vendors (Harvey for legal, Hippocratic AI for
healthcare, Sierra for customer experience) sit between \emph{open
standards} and \emph{production hybrid}: the bundle is the commercial
unit; the underlying model is mostly the same one a competitor would
use.

\subsection{8. The two-loop design
space}\label{the-two-loop-design-space}

The fourteen composite-two-loop systems sit in a 3×3 matrix indexed by
gate type on each loop.

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Loop-2 Auto-gated}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Loop-2 Human-gated}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Loop-2 None}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Loop-1 Auto-gated} & NanoResearch, SkillRL, AlphaEvolve, ADAS,
DGM (primary), Voyager, SAGE, EvolveR, CoEvoSkills, AutoAgent,
AutoManual & SkillForge, CASCADE, DGM (sandbox), Sierra OS 2.0,
Cognition Devin & MAE*, MetaGen*, MetaReflection*, CLIN* \\
\textbf{Loop-1 Human-gated} & \textbf{{[}empty by engineering logic{]}}
& OpenCore & \emph{(none)} \\
\textbf{Loop-1 None} & ExpeL (weak), Braintrust*, LangSmith Hub*,
Vellum* & Anthropic skills repo*, DSPy*, TextGrad* & Self-Refine*,
ReAct*, Mem0* \\
\end{longtable}
}

\emph{Asterisks fail at least one composite-two-loop criterion; included
for contrast.}

The matrix has four cells of interest. \textbf{{[}Auto × Auto{]}} is
most populated (eleven systems). Every member has a machine-evaluated
decision criterion on both loops. \textbf{{[}Auto × Human{]}} is the
production-viable cell --- Loop 1 fires at machine speed under an
algorithmic criterion, Loop 2 requires human approval before
propagation. \textbf{{[}Human × Human{]}} contains only OpenCore: the
dream → delta → PR machinery is automated throughout, but both gates ---
\texttt{core-update-gate.md} for Loop 1 PR review and the upstream-CORE
PR for Loop 2 --- are reviewer-judgment checklists rather than
algorithmic criteria. \textbf{{[}Human × Auto{]}} is empty by
engineering logic; if humans author candidates manually, automated
promotion offers little leverage. A speculative occupant would be
``expert curation + auto-distribution'' (expert-authored skills
automatically gated through a held-out eval before shipping), but no
published example exists.

The \textbf{{[}None × Auto-gated{]}} cell holds the governance-first
prompts-as-code systems (Braintrust, LangSmith Hub, Vellum). They have a
real Loop 2 --- eval-gated CI blocks promotion on regression --- but no
Loop 1 in the patch-plastic sense, because the candidate artifacts are
authored by engineers at their desks rather than triggered by session
feedback. The asterisks mark this: governance without inner-loop
adaptation. The cell is informative precisely because it shows what
survives when only the outer loop is automated.

OpenCore deserves one paragraph here because its ML-mapping is unusually
clean and the four-layer architecture is unique in the corpus. The dream
→ 5-dream rule is the only published \(n > 1\) minibatch in the survey:
each ``dream'' is a candidate gradient sample at \(n = 1\), and the
5-dream threshold defers the fork update until five independent dreams
converge --- directly mirroring mini-batch gradient descent at batch
size 5. The four-layer fork (USER → PROJECT → STACK → CORE) is the only
explicit deep-vs-surface hierarchy in the corpus, with CORE as the
slow-moving universal layer and USER as the fast-adapting per-instance
layer. Cross-project drift upstream-promoted to CORE is federated
averaging across user/project shards, gated by human PR review. The key
observation here is that OpenCore demonstrates \emph{cross-organisation}
federated aggregation --- the only surveyed system to do so explicitly
--- and that the gates are reviewer-judgment because the design is
targeted at solo and small-team workflows where the marginal cost of a
bad auto-merge exceeds the marginal cost of human latency.

Two cross-cutting findings sharpen the picture. \textbf{Noise reduction
via batching shows up only under human gates} in the surveyed corpus:
OpenCore's 5-dream rule is the only published \(n > 1\) minibatch, and
it sits in {[}Human × Human{]}. The auto-gated systems all fire at
\(n = 1\) per generation step, absorbing noise statistically via the
algorithmic gate. Whether the implicit minibatch in OpenCore reduces
residual error against an \(n = 1\) baseline is untested; we predict it
does. \textbf{Multi-layer hierarchies are under-explored}: only OpenCore
and AlphaEvolve demonstrate explicit depth. The other twelve
composite-two-loop systems treat their artifact pool as flat. Adding a
STACK-style intermediate layer to NanoResearch or a multi-archetype
scope to Anthropic's skills repository is a tractable engineering
direction.

\subsection{9. The empty corner}\label{the-empty-corner}

Four properties define what we call \textbf{enterprise full-auto}:

\begin{itemize}
\tightlist
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \tightlist
  \item
    \textbf{Auto-gated Loop 1} --- the commit criterion is an
    algorithmic threshold.
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{1}
  \tightlist
  \item
    \textbf{Multi-tenant cross-org promotion} --- Loop 2 routes
    artifacts between organisations, not just within one.
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{2}
  \tightlist
  \item
    \textbf{Versioned lineage with RBAC} --- promoted artifacts carry
    semantic versions, audit trails, and access-control policies.
  \end{enumerate}
\item
  \begin{enumerate}
  \def\labelenumi{(\alph{enumi})}
  \setcounter{enumi}{3}
  \tightlist
  \item
    \textbf{Rollback} --- a bad promotion can be reverted without
    redeploy.
  \end{enumerate}
\end{itemize}

No surveyed system has all four. Closest approaches partition the
requirements: \textbf{AlphaEvolve} has (a) and partial (b) within a
single organisation; \textbf{NanoResearch} has (a) at the research
level; \textbf{SkillForge} has (a) and (b) within one enterprise,
partial (c) via VFS commits, undocumented (d); \textbf{OpenCore} has
(b), (c), and (d) under reviewer-judgment gates but not (a);
\textbf{Sierra Agent OS 2.0} has (b) and partial (a); \textbf{Vellum /
Braintrust / LangSmith Hub} have (c) and (d) explicitly but no (a) or
(b) in the strict sense.

The empty corner is concrete in ML terms: a \emph{DP-FedAvg-style
cross-tenant aggregator with an automated eval gate at the scaffold
layer}. Differential-privacy federated averaging (DP-FedAvg) is the
closest analogue in the federated-learning literature --- gradient
updates aggregated across distributed shards with privacy-preserving
noise and an aggregation gate. At the scaffold layer this would mean:
per-tenant artifact pools with tenant-tagged provenance, aggregation
across tenants with operator-tunable privacy bounds, an automated eval
gate (held-out task suite per tenant) blocking promotion when patch loss
does not decrease, versioned lineage with RBAC carrying tenant access
policies forward through the promotion, and a rollback envelope on every
promoted artifact. Naming the corner converts a vague gap into a
specific four-property audit criterion; the implementor's safety
requirements are deferred to follow-on practitioner work. It is no
longer ``no system closes both loops'' --- too coarse --- but a specific
four-property combination that the survey can audit any future system
against.

Whether this corner \emph{should} be filled is deployment-dependent. In
safety-critical domains --- medical decision support, legal advice,
financial advice --- reviewer-judgment gates may be a design feature
rather than a defect, and the human latency is the point. The narrower
survey claim is what matters: no public system currently combines
automated local scaffold evolution with governed cross-organisation
promotion, versioned lineage with RBAC, and rollback. Naming the corner
converts a vague gap into a specific four-property audit criterion that
any future system can be measured against. The shortest demonstrable
path appears to build on AlphaEvolve's cascade evaluator + MAP-Elites
archive, extending federation from within-DeepMind multi-target to
cross-customer multi-tenant --- but the architectural surface, not the
timeline, is what this paper claims.

\subsubsection{9.1 Patch-complete systems}\label{patch-complete-systems}

The two-tier architecture also changes the practical meaning of
``general'' capability. A frontier model need not be globally general
for the deployed system to become effectively general inside a bounded
patch. Once a localhost scaffold contains the patch's rules, tools,
memories, workflows, evals, and promotion gates, the model--scaffold
composite can cover the economically relevant task distribution of that
patch with broad competence. We call this condition
\textbf{patch-completeness}: the system is not generally intelligent
over the world, but it is operationally general over the bounded world
it inhabits.

Patch-completeness also reframes what many deployments experience as
``general intelligence.'' In practice, the deployed system is often
closer to a language-mediated local learner than to a universal
reasoner. The frontier model supplies the broad linguistic interface and
cross-domain prior; the scaffold supplies the patch-specific feature
engineering, memory, tools, rules, eval criteria, workflow state, and
loss surface. Once those local coordinates are rich enough, the
model--scaffold composite behaves less like a naked foundation model and
more like a fitted decision system over the bounded world of the patch.

This is the old lesson of supervised learning returning through a
language interface. A workflow-specific agent resembles a local function
approximator: it maps recurring situations in the patch to acceptable
actions under a locally defined loss. The difference is that the
features are no longer hand-built columns in a tabular model; they are
retrieved documents, project rules, tool states, memory entries, tests,
schemas, reviewer expectations, and governance gates. The intelligence
appears general because the patch itself supplies the missing structure.

The consequence is that ``AGI'' is the wrong first unit for many
production systems. The operational question is not whether one model
has mastered the world, but whether a model--scaffold system has become
patch-complete: broad enough to cover the economically relevant task
distribution inside one bounded world, and governed enough not to
mistake that local fit for global competence.

This is why many production systems already feel further along than
model-only benchmarks imply. A coding agent inside one repository ---
with project rules, tests, CI, documentation, tool access, issue
history, and PR review --- may be patch-complete for that repository. A
legal agent inside a firm's due-diligence workflow, with vetted document
stores, clause libraries, playbooks, and review gates, may be
patch-complete for that workflow. The competence is not located in the
model alone. It is distributed across the frontier prior and the
localhost scaffold.

Patch-completeness is not global generality. It is local generality
under boundary conditions. The distinction matters because the failure
mode changes: a patch-complete system can appear broadly competent while
silently depending on local conventions, hidden evaluator expectations,
and scaffold assumptions that do not transfer. Moving it to a
neighbouring patch without re-measuring the residual catalogue is
therefore a patch-shift event, not ordinary deployment. The governance
problem is not merely how to make scaffolds learn but \emph{how to know
where their local generality ends}. If patch-complete systems exist ---
and many deployed systems likely already are --- then scaffold
governance matters because local competence can become powerful before
global robustness exists. §10 catalogues the failure surface that newly
powerful local systems exhibit.

\subsection{10. A new failure surface}\label{a-new-failure-surface}

Patch-plasticity introduces failure modes weight-only systems do not
have. We name them here; the full taxonomy (nine categories A--I with
severity ratings and counter-principles P1--P8) is deferred to follow-on
practitioner work.

\textbf{Scaffold bloat} is the dominant near-term risk. IFScale
(Jaroslawicz et al., 2025) reports a 35× cost increase as instruction
count scales from 10 to 250, with frontier-model accuracy collapsing to
68\% at 500 instructions. Lost-in-the-middle (Liu et al., 2024) supplies
the mechanism. \textbf{Poisoned memory} is the dominant security risk:
PoisonedRAG (Zou et al., 2024) reports 90\%+ attack success rate against
standard RAG; Memory Control Flow Attacks (Xu et al., 2026) report 100\%
persistence after injection. \textbf{Prompt injection via tools and MCP}
is the dominant cross-vendor attack surface, with CVE-2025-54136
(``MCPoison'') demonstrating a public MCP-server compromise pathway.
\textbf{Eval fragility} is the dominant ecosystem risk: the Leaderboard
Illusion (Lin et al., 2025) documents up to 112\% inflation on common
benchmarks under optimisation-on-test-set effects; eval suites are
themselves patch-plastic artifacts and must survive Goodhart's Law.
\textbf{Coordination failures} in multi-agent systems are documented in
MAST (Cemri et al., 2025) --- 14 distinct failure modes with
\(\kappa = 0.88\) agreement, dominantly role-confusion / context-loss /
coordination rather than prompt quality. \textbf{Overfitting to patch}
is structural: scaffolds tuned to a deployment over-fit it, and the
framework predicts but does not measure how badly. \textbf{Plasticity
loss at the scaffold level} is the slow-moving risk: context rot,
instruction conflict, memory pollution, and tool-version drift produce a
scaffold-level analogue of weight-level plasticity loss (Dohare et
al.~2024, Lyle et al.~2023). Scaffolds age; a six-month-old
\texttt{CLAUDE.md} is often worse than a clean rebuild, and no surveyed
system treats pruning or expiration as a first-class operation.

Taken together: the patch-plastic surface is real, quantified by
2024--2026 work, and largely unmitigated in production. Follow-on
practitioner work walks the mitigation principles in detail.

\subsection{11. Conclusion}\label{conclusion}

The field did not wait for frontier weights to become perfectly
reliable. It built a second learning system around them. That second
system is made of markdown files, skills, memories, tools, orchestration
graphs, eval suites, PRs, version histories, and rollback buttons. It
looks like engineering clutter until viewed as a scaffold parameter
vector under update. Then the clutter resolves into architecture:
production LLMs learn locally, outside the weights.

The mechanisms were already there. A Claude Skill has been a procedural
weight since Anthropic shipped progressive disclosure. A Cursor Rule has
been a local parameter since \texttt{.cursor/rules/} became a directory.
A PR against \texttt{agents.md} has been a promotion gate since the
Linux Foundation took stewardship. What changed in 2025--2026 is that
the \emph{combination} --- these substrates wrapping a frozen frontier
model and updating from deployment signal --- became dense enough across
vendors that the convergence reads as architecture rather than
coincidence. Once you name it, the empty corners of the design space
become concrete engineering targets. The most consequential one is
enterprise full-auto: Auto-gated Loop 1 plus governed multi-tenant
cross-org Loop 2, with versioned lineage, RBAC, and rollback. In ML
language, a DP-FedAvg-style cross-tenant aggregator with an automated
eval gate at the scaffold layer. Whether it should be filled is
deployment-dependent; that it has not been is the survey's headline
finding.

The five surveyed absences of §6.3 --- failure-triggered memory rare in
production, cross-customer skill promotion absent, memory versioning
unsolved, scaffold-level curriculum unspecified, Loop-2 overfitting
detection missing --- are each a Paper-4 target. None of these is novel
as a complaint; each is novel as a \emph{gap the architecture makes
visible}. The patch-completeness observation of §9.1 sharpens the
urgency: local competence is becoming powerful enough to require
governance attention before global robustness has been demonstrated.

The deepest connection --- back to Paper 2 --- is structural.
Patch-plasticity is plastic because residual errors are clustered; if
mode discovery were not slow (Postulate 1 in \emph{Architecture of
Errors}), patches would not generalise across deployments and the
architecture would collapse. The survey is consistent with the postulate
at the system level: deployments do generalise enough of the patch to
make scaffold investment pay off. Whether this remains true at agentic,
scientific, and long-horizon scales is the empirical test the next paper
should design.

Practical generality therefore arrives locally first: not as a universal
mind, but as patch-complete model--scaffold systems whose competence is
fitted to bounded worlds. The frontier model generalises. The localhost
scaffold specialises. Reliability comes from governing that
specialisation. The mechanisms were already there --- markdown files,
skills, memories, tools, orchestration graphs, eval suites, PRs, version
histories, rollback buttons --- but they read as engineering clutter
until viewed as a scaffold parameter vector under update. Then the
clutter resolves into architecture: production LLMs learn locally,
outside the weights, because that is the only place patch-specific
adaptation can safely happen.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Appendix A. Representative survey
rows}\label{appendix-a.-representative-survey-rows}

The full per-system spreadsheet is \texttt{p3\_master\_scores.csv} in
the repository. The 30 rows below are a representative spot-check
covering all six substrates, both research and production, and all five
clusters of §6.2. \emph{Loop 1} and \emph{Loop 2} columns apply to
systems satisfying the composite-two-loop criterion; \emph{n/a} means
the loop is absent for that system. The \emph{PP} column is the
integrated patch-plasticity score (0--5) used in the survey; the
\emph{Evidence} column abbreviates the strongest available source tier
(T1 peer-reviewed; T2 arXiv preprint; T3 production claim; T4
open-source artifact; T5 industry blog).

\begingroup\small

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 14\tabcolsep) * \real{0.1212}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 14\tabcolsep) * \real{0.0404}}
  >{\raggedright\arraybackslash}p{(\linewidth - 14\tabcolsep) * \real{0.1111}}
  >{\raggedright\arraybackslash}p{(\linewidth - 14\tabcolsep) * \real{0.0707}}
  >{\raggedright\arraybackslash}p{(\linewidth - 14\tabcolsep) * \real{0.0707}}
  >{\raggedleft\arraybackslash}p{(\linewidth - 14\tabcolsep) * \real{0.0404}}
  >{\raggedright\arraybackslash}p{(\linewidth - 14\tabcolsep) * \real{0.1818}}
  >{\raggedright\arraybackslash}p{(\linewidth - 14\tabcolsep) * \real{0.3636}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
System
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Year
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Substrate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Loop 1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Loop 2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PP
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evidence
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Why included
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
NanoResearch & 2026 & Integrated & Auto & Auto & 5 & T2
(arXiv:2605.10813) & Tri-level Skills/Memory/Policy co-evolution;
cleanest full-auto exemplar \\
AutoAgent & 2026 & Integrated & Auto & Auto & 5 & T2 (arXiv:2603.09716)
& Dual-cycle Execution+Evolution + elastic memory orchestration \\
SkillRL & 2026 & Integrated & Auto & Auto & 5 & T2 (arXiv:2602.08234) &
Recursive skill-augmented RL with \(\text{SR}<0.4\) threshold gate \\
AlphaEvolve & 2025 & Integrated & Auto & Auto & 4 & T2/T3
(arXiv:2506.13131) & Production proof-point: Borg +0.7\%, Gemini kernel
+23\%, FlashAttention +32.5\% \\
DGM (primary) & 2025 & Integrated & Auto & Auto & 4 & T2
(arXiv:2505.22954) & Recursive self-modifying code; SWE-bench
20\%→50\% \\
ADAS & 2024 & Integrated & Auto & Auto & 4 & T2 (arXiv:2408.08435) &
Meta-agent search; Turing-complete agent representation \\
SkillForge & 2026 & Integrated & Auto & Human & 4 & T2
(arXiv:2604.08618) & Production hybrid exemplar; LLM-judge
\textgreater90\% + VFS versioning \\
CASCADE & 2025 & Integrated & Auto & Human & 4.5 & T2 (arXiv:2512.23880)
& Largest published ablation gain in corpus: +57.9 pp on
SciSkillBench \\
OpenCore & 2025 & Integrated & Human & Human & 4 & T4
(\url{github.com/sibmike/opencore}) & Only {[}Human × Human{]} system
with explicit cross-organisation federation \\
Voyager & 2023 & Integrated & Auto & Auto & 4 & T2 (arXiv:2305.16291) &
Foundational skill-library result; 3.3× items, 15.3× tech-tree
milestones \\
AGENTS.md & 2025 & S1 Instructions & n/a & n/a & 2 & T3/T4 (agents.md) &
Cross-vendor standard; 60k+ repos; Lulla 2026 measures −28.6\%
runtime \\
Claude Code CLAUDE.md & 2025 & S1 Instructions & n/a & n/a & 4 & T3/T4
(docs.anthropic.com) & Four scoping levels including org-IT policy +
auto-memory layer; ceiling for S1 \\
Cursor Rules & 2024 & S1 Instructions & n/a & n/a & 3 & T3/T4
(\url{docs.cursor.com/rules}) & \texttt{.cursor/rules/*.mdc} multi-scope
+ MCP attach; git-tracked \\
Anthropic Agent Skills & 2025 & S2 Skills & n/a & Human & 3 & T3
(\url{anthropic.com/engineering}) & Progressive disclosure spec;
LangChain replication 29\%→95\% pass-rate \\
SAGE & 2025 & S2 Skills & Auto & Auto & 5 & T2 (arXiv:2512.17102) &
+8.9\% SGC, 26\% fewer steps, 59\% fewer tokens on AppWorld \\
COSPLAY & 2026 & S2 Skills & Auto & Auto & 5 & T2 (arXiv:2604.20987) &
Boundary proposal + segmentation; only S2 = 5 with contracts \\
Reflexion & 2023 & S3 Memory & Auto & n/a & 4 & T1 (NeurIPS 2023;
arXiv:2303.11366) & Canonical failure-triggered episodic memory; +8\%
HotpotQA \\
Generative Agents & 2023 & S3 Memory & Auto & n/a & 4 & T1 (UIST 2023;
arXiv:2304.03442) & Reflection + importance memory primitive \\
Governed Collaborative Memory (SSGM) & 2026 & S3 Memory & Auto & Auto &
5 & T2 (arXiv:2605.04264) & Only memory system with full provenance +
versioning + correction \\
MCP & 2024 & S4 Tools & n/a & n/a & 5 & T3 (modelcontextprotocol.io) &
Cross-vendor; 9,400+ servers; 78\% enterprise; \textasciitilde97M SDK
downloads/month \\
Harvey & 2026 & S4 Tools & n/a & n/a & 5 & T3 (harvey.ai) & 400K
queries/day; 18,000+ workflows; 200+ legal data sources \\
Hippocratic AI & 2026 & S4 Tools & n/a & n/a & 5 & T3
(hippocraticai.com) & \$3.5B valuation; 30\% readmission reduction;
360\% care-capacity boost \\
Multi-Agent Evolve (MAE) & 2025 & S5 Orchestration & Auto & n/a & 5 & T2
(arXiv:2510.23595) & RL co-evolution of Proposer/Solver/Judge
population \\
Memento-Skills & 2026 & S5 Orchestration & Auto & Auto & 5 & T2
(arXiv:2604.02460) & Unit-test-gated skill promotion within multi-agent
topology \\
Cemri et al.~(AG2) & 2025 & S5 Orchestration & n/a & n/a & 3 & T2
(arXiv:2503.13657) & Specialisation +4.5 pp on GSM-Plus (\(p = 0.03\));
MAST 14 failure modes \\
Braintrust & 2024 & S6 Governance & n/a & Auto & 5 & T3 (braintrust.dev)
& Prompts-as-code: immutable commits, eval-gated PR merge, sub-5 min
rollback \\
Vellum & 2024 & S6 Governance & n/a & Auto & 5 & T3 (vellum.ai) & Same
governance pattern; production deployment with eval thresholds \\
LangSmith Hub & 2024 & S6 Governance & n/a & Auto & 5 & T3
(\url{smith.langchain.com}) & Prompt repository with environment
promotion and eval blocking \\
Self-Refine* & 2023 & (contrast) & n/a & n/a & 1 & T2 (arXiv:2303.17651)
& Fails the patch-plastic discriminator: in-context iteration only \\
Mem0* & 2024 & (contrast) & n/a & n/a & 2 & T3/T4 (mem0.ai) & Fails Loop
2: per-user memory with no cross-context promotion path \\
\end{longtable}
}

\endgroup

\emph{Asterisks mark contrast systems included for definitional
clarity.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{References}\label{references}

ACE authors. (2025). Agentic Context Engineering: Evolving playbooks via
generation, reflection, and curation. \emph{arXiv preprint
arXiv:2510.04618}.

AGENTS.md. (2025). AGENTS.md: A cross-vendor open standard for agent
instructions. Linux Foundation. \url{https://agents.md}

Anthropic. (2024). Model Context Protocol specification.
\url{https://modelcontextprotocol.io}

Arbuzov, M. L., Shvets, A. A., \& Bei, S. (2025). Beyond exponential
decay: Rethinking error accumulation in large language models.
\emph{arXiv preprint arXiv:2505.24187}.

Arbuzov, M. L., Shvets, A. A., \& Bei, S. (2026). The architecture of
errors: Logarithmic mode discovery and polylogarithmic intervention
budgets for long-context LLM reliability. \emph{arXiv preprint}
(forthcoming).

Bao, M., et al.~(2024). AutoManual: Constructing instruction manuals by
LLM agents via interactive environmental learning. \emph{NeurIPS 2024}.
arXiv:2405.16247.

Cao, Y., et al.~(2025). Mobile-Agent-E: Self-evolving mobile assistant
for complex tasks. \emph{arXiv preprint arXiv:2501.11733}.

Cemri, M., et al.~(2025). MAST: A taxonomy of failure modes in
multi-agent LLM systems. \emph{arXiv preprint arXiv:2503.13657}.

Chen, S., et al.~(2024). Agent S: An open agentic framework that uses
computers like a human. \emph{arXiv preprint arXiv:2410.08164}.

Chen, S., et al.~(2025). Agent S2: Compositional open agentic framework.
\emph{arXiv preprint arXiv:2504.00906}.

Chen, W., et al.~(2026). Governed collaborative memory / SSGM.
\emph{arXiv preprint arXiv:2605.04264}.

Dohare, S., et al.~(2024). Loss of plasticity in deep continual
learning. \emph{Nature}, 632(8026), 768--774.

Du, P. (2026). Memory for autonomous LLM agents: Mechanisms, evaluation,
and emerging frontiers. \emph{arXiv preprint arXiv:2603.07670}.

Fang, J., Peng, Y., Zhang, X., et al.~(2025). A comprehensive survey of
self-evolving AI agents. \emph{arXiv preprint arXiv:2508.07407}.

Gao, H.-A., Geng, J., Hua, W., et al.~(2026). A survey of self-evolving
agents: What, when, how. \emph{Transactions on Machine Learning
Research}. arXiv:2507.21046.

Harada, T., et al.~(2025). The curse of instructions: Compliance
collapse under instruction-set growth. \emph{ICLR 2025}.

Harvey. (2026). Harvey product overview. \url{https://www.harvey.ai}

Hippocratic AI. (2026). Polaris clinical outcome evidence.
\url{https://www.hippocraticai.com}

Hu, S., Lu, C., \& Clune, J. (2024). Automated design of agentic systems
(ADAS). \emph{arXiv preprint arXiv:2408.08435}.

Huang, T., et al.~(2025). CASCADE: A collaborative and adaptive science
agent with skill distillation and evolution. \emph{arXiv preprint
arXiv:2512.23880}.

Jaroslawicz, K., et al.~(2025). IFScale: Instruction-following at scale.
\emph{arXiv preprint arXiv:2507.11538}.

Jiang, P., Lin, J., Shi, Z., et al.~(2025). Adaptation of agentic AI: A
survey of post-training, memory, and skills. \emph{arXiv preprint
arXiv:2512.16301}.

Lin, T., et al.~(2025). The leaderboard illusion: Quantifying
optimisation-on-test-set effects in LLM benchmarks. \emph{arXiv
preprint}.

Liu, N. F., et al.~(2024). Lost in the middle: How language models use
long contexts. \emph{TACL}.

Liu, Z., et al.~(2025). Multi-agent evolve: RL for joint
Proposer/Solver/Judge optimisation. \emph{arXiv preprint
arXiv:2510.23595}.

Lulla, A., et al.~(2026). Empirical evaluation of AGENTS.md as a
scaffold standard. \emph{arXiv preprint arXiv:2601.20404}.

Lyle, C., et al.~(2023). Understanding plasticity in neural networks.
\emph{arXiv preprint arXiv:2303.07507}.

Newell, A., et al.~(1990). \emph{Unified theories of cognition}. Harvard
University Press.

OpenCore. (2025). OpenCore: Open-source LLM-assisted coding practices.
\url{https://github.com/sibmike/opencore}. MIT License.

Park, J. S., et al.~(2023). Generative agents: Interactive simulacra of
human behavior. \emph{UIST 2023}. arXiv:2304.03442.

Press, O., et al.~(2022). Measuring and narrowing the compositionality
gap in language models. \emph{arXiv preprint arXiv:2210.03350}.

Shinn, N., et al.~(2023). Reflexion: Language agents with verbal
reinforcement learning. \emph{NeurIPS 2023}. arXiv:2303.11366.

Sierra. (2024). Sierra Agent OS. \url{https://sierra.ai}

Sierra. (2025). Agent OS 2.0: From answers to memory and action.
\url{https://sierra.ai/blog/agent-os-2-0}

SkillForge authors. (2026). SkillForge: Forging domain-specific
self-evolving agent skills in cloud technical support. \emph{arXiv
preprint arXiv:2604.08618}.

SkillRL authors. (2026). SkillRL: Evolving agents via recursive
skill-augmented reinforcement learning. \emph{arXiv preprint
arXiv:2602.08234}.

Sumers, T. R., Yao, S., Narasimhan, K., \& Griffiths, T. L. (2023).
Cognitive architectures for language agents (CoALA). \emph{Transactions
on Machine Learning Research}. arXiv:2309.02427.

Tan, W., et al.~(2024). Cradle: Empowering foundation agents towards
general computer control. \emph{arXiv preprint arXiv:2403.03186}.

Tran, S., \& Kiela, D. (2026). Memento-Skills: Unit-test-gated skill
promotion for multi-agent systems. \emph{arXiv preprint
arXiv:2604.02460}.

Wang, G., et al.~(2023). Voyager: An open-ended embodied agent with
large language models. \emph{arXiv preprint arXiv:2305.16291}.

Wang, X., et al.~(2026). AutoAgent: Evolving cognition and elastic
memory orchestration for adaptive agents. \emph{arXiv preprint
arXiv:2603.09716}.

Wu, R., et al.~(2025). EvolveR: Self-evolving LLM agents through an
experience-driven lifecycle. \emph{arXiv preprint arXiv:2510.16079}.

Wu, Z., et al.~(2024). OS-Copilot: Towards generalist computer agents
with self-improvement. \emph{arXiv preprint arXiv:2402.07456}.

Xu, J., et al.~(2026). NanoResearch: Co-evolving skills, memory, and
policy for personalized research automation. \emph{arXiv preprint
arXiv:2605.10813}.

Xu, Z., et al.~(2026). Memory control flow attacks against large
language model agents. \emph{arXiv preprint} (MCFA).

Zhang, H., et al.~(2026). CoEvoSkills: Self-evolving agent skills via
co-evolutionary verification. \emph{arXiv preprint arXiv:2604.01687}.

Zhang, J., et al.~(2025). Darwin Gödel machine: Open-ended evolution of
self-improving agents. \emph{arXiv preprint arXiv:2505.22954}.

Zhou, Y., Shu, W., Su, Y., et al.~(2026). A comprehensive survey on
agent skills: Taxonomy, techniques, and applications. \emph{arXiv
preprint arXiv:2605.07358}.

Zhou, Y., et al.~(2025). SAGE: Skill-augmented GRPO for self-evolution.
\emph{arXiv preprint arXiv:2512.17102}.

Zou, W., et al.~(2024). PoisonedRAG: Knowledge corruption attacks on
retrieval-augmented generation. \emph{arXiv preprint}.

(DeepMind AlphaEvolve authors). (2025). AlphaEvolve: A coding agent for
scientific and algorithmic discovery. \emph{arXiv preprint
arXiv:2506.13131}.

\end{document}
