% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Cambria}
  \setmonofont[]{Consolas}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The Architecture of Errors: Logarithmic Mode Discovery and Polylogarithmic Intervention Budgets for Long-Context LLM Reliability},
  pdfauthor={Mikhail L. Arbuzov; Alexey A. Shvets; Sisong Bei},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{The Architecture of Errors: Logarithmic Mode Discovery and
Polylogarithmic Intervention Budgets for Long-Context LLM Reliability}
\author{Mikhail L. Arbuzov \and Alexey A. Shvets \and Sisong Bei}
\date{}

\begin{document}
\maketitle

\textbf{Abstract}

Our previous work (Arbuzov, Shvets, \& Bei, 2025) argued that LLM
reliability does not decay exponentially with output length because
errors concentrate at sparse ``key tokens'' rather than spreading
uniformly across the sequence. This paper extends that framework along
an orthogonal axis: LLM errors are not only sparse, they are
\emph{repetitive}. The same handful of failure patterns recur across
models, datasets, and domains. Inside the small set of key tokens, a
latent fraction \(\beta\) produce \emph{hard} failures; those hard
failures cluster into a finite catalogue of recurring modes whose count
grows slowly with the number of observed failures. Under an empirical
postulate of logarithmic mode discovery, anchored by three published
taxonomies (ErrorAtlas, MWPES, HumanEval-error-categorisation), the
intervention budget required to bound \emph{per-hard-token} residual
error scales polylogarithmically in sequence length. Sequence-level
reliability targets are strictly tighter and approach full-catalogue
coverage as the number of hard decisions grows --- an asymmetry we
surface explicitly rather than burying. Twelve intervention categories
from prior work, plus 27 GOLD-tier capability-elimination citations
across six axes, are consistent with a small library on the order of
tens of interventions covering the head of the failure distribution in
the per-hard-token regime. We test the polylog conclusion under four
cluster-count laws (logarithmic discovery, Heaps at
\(b \in \{0.3, 0.5, 0.6\}\), saturating exponential) and find it robust
across the family, though the specific doubly-logarithmic rate is the
optimistic special case. Prominent steep-decay counter-evidence (Dziri
et al., 2023; Kuratov et al., 2024; METR, 2025), re-audited here, decays
over variables other than raw token length --- compositional graph size,
fact count, log-time horizon. This relocates rather than dissolves the
practical worry: the framework identifies which interventions help
(capability provisioning along the actual decay axis) and which do not,
but does not claim the regimes where \(k_{\text{hard}}\) grows fast are
thereby made easy.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Introduction}\label{introduction}

The standard worry about long-context autoregressive generation is
exponential. If every token has independent error probability \(e\),
then after \(n\) tokens the chance of a fully correct output is
\((1-e)^n\), which collapses to zero for any nontrivial \(e\) and large
enough \(n\) (LeCun, 2023; Dziri et al., 2023). In our previous paper
(Arbuzov, Shvets, \& Bei, 2025) we argued that this worry is misplaced
because errors are not distributed uniformly across tokens. Empirical
work (Fang et al., 2024) shows that only 5--10\% of tokens are ``key''
--- genuinely dependent on long-range context --- while the remaining
majority become nearly deterministic once enough context accumulates.
Replacing \((1-e)^n\) with the two-rate model

\[
P(\text{correct}) = (1-e_{\text{key}})^{k} \cdot (1-e_{\text{non}})^{n-k}, \qquad k \ll n,
\]

with \(k\) scaling sublinearly in \(n\), recovers the observed
long-context coherence of modern LLMs and converts the reliability
question from ``how does \(n\) grow?'' to ``how does \(k\) grow?''.

\subsubsection{1.1 From sparse tokens to recurring
patterns}\label{from-sparse-tokens-to-recurring-patterns}

This paper takes the next step. Sparsity tells us \emph{where} errors
live. The natural follow-up question is \emph{what} they are. Recent
failure-mode atlases supply a striking empirical answer: errors are not
only sparse, they are also \emph{repetitive}. The ErrorAtlas project
(Ashury-Tahan et al., 2026), covering 83 models across 35 datasets,
finds that the top five of seventeen named failure categories account
for 53.6\% of all observed errors and the top ten for over 80\%. In
code, two error types (\texttt{AssertionError} and \texttt{NameError})
cover 86.35\% of failures on HumanEval across 14 LLMs, a result
replicated across 23 models on HumanEval Pro and MBPP Pro (Zhang et al.,
2024). In math, the top four categories per model capture 45--55\% of
errors across 15 LLMs and four benchmark datasets, with the same
hierarchy across models (Sun et al., 2025). The same picture appears in
multi-hop QA (Weakest Link Law authors, 2026), agentic tool use (MAST
authors, 2025), and retrieval-augmented generation (Acurai authors,
2024).

These observations suggest a third layer in the reliability
architecture, complementing key-token sparsity (Layer 1) and the
within-key stratified-manifold structure (Layer 2 in Arbuzov et al.,
2025). Within the small set of key tokens, only a fraction \(\beta\)
produce \emph{hard} failures, and those hard failures cluster into a
finite catalogue of recurring modes whose size grows much more slowly
than the number of observed events.

\subsubsection{1.2 The intervention
budget}\label{the-intervention-budget}

If failure modes form a small, slowly-growing catalogue, then targeted
interventions become a viable engineering strategy. Each recurring mode
admits a specific countermeasure: arithmetic errors fall to a Python
interpreter (Gao et al., 2022; Chen et al., 2022; Gou et al., 2023);
format violations to constrained decoding (Suresh et al., 2025; SLOT
authors, 2025); hallucinations to retrieval augmentation, with one paper
reporting 100\% elimination on RAGTruth (Acurai authors, 2024);
over-refusals to preference optimisation (Karaman et al., 2024);
tool-call errors to structured uncertainty (Suri et al., 2025). We
collect twelve such categories with quantitative ``one cluster, one
intervention'' results in Section 4.2.

The natural question becomes: how many interventions are enough? If new
failure modes follow canonical Heaps-style growth (Manning et al.,
2008), the optimistic doubly-logarithmic rate fails, but the
intervention budget can still remain polylogarithmic when the number of
hard decisions grows sublinearly. The practical question is therefore
the exponent and constants, not whether the program is hopeless. The
data also suggest the more optimistic case: error-taxonomy work plateaus
quickly, with most domains stabilising at 10--50 recurring modes even at
large corpus sizes. We treat this as an \emph{empirical postulate of
logarithmic mode discovery} (Section 3.2) rather than as a derived
consequence --- Zipfian rank-frequency does \emph{not} imply logarithmic
cluster growth in the strict mathematical sense; Heaps' law is the
correct type--token consequence of Zipf-distributed events, and it is
power-law, not logarithmic. Under our postulate, the required
intervention budget scales polylogarithmically in sequence length
(Proposition 1, Section 3.4).

\subsubsection{1.3 Contributions}\label{contributions}

The paper does five things. It introduces a three-layer formal framework
--- sparsity (\(\alpha\)), stratification (\(\beta\)), and bounded mode
catalogue (\(|C|\)) --- that refines our previous two-rate model by
partitioning key tokens into easy and hard subsets and adding a
recurring-pattern catalogue. It posits, and empirically calibrates,
logarithmic mode discovery: anchored against ErrorAtlas, HumanEval, and
MWPES, a single mode-rate parameter \(\sigma \approx 2\) fits all three.
It proves a polylog bound on the per-hard-token intervention library
size, \(m \geq |C|^{1 - \varepsilon / e_{\text{hard}}}\), and shows that
under \(k \sim \log n\) the budget grows as
\(O\big((\log\log n)^{c}\big)\) with \(c < 1\); the
\emph{sequence}-level analogue is strictly tighter and approaches
full-catalogue coverage. It synthesises evidence for the three
load-bearing claims --- clustering (Section 4.1), cluster-orthogonal
interventions (Section 4.2), and sublinear length scaling (Section 4.3)
--- drawing on approximately sixty prior published results, including a
six-axis capability-elimination harvest with 27 GOLD-tier citations. And
it re-audits the most-cited steep-decay results (Dziri et al., 2023;
Kuratov et al., 2024; METR, 2025; Wan et al., 2025), showing that each
decays over a variable other than raw token length and so is compatible
with the sublinear-in-\(n\) thesis.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Related Work}\label{related-work}

\subsubsection{2.1 Failure-mode
taxonomies}\label{failure-mode-taxonomies}

The systematic cataloguing of LLM failures has reached maturity across
several domains in the last two years. ErrorAtlas (Ashury-Tahan et al.,
2026) provides the largest cross-domain taxonomy to date: 83 models
evaluated on 35 datasets yield approximately 7,000 sampled failures
organised into 17 named categories, with the top-5/top-10 cumulative
coverage figures cited above. MWPES-300K (Sun et al., 2025) takes the
math domain deeper, with 304,865 categorised errors from 15 LLMs across
four math word-problem datasets and a strict per-model decomposition
that shows the same top-4 categories dominating across models --- a
striking dataset-invariance result that we use heavily in Section 4.1.
For code, Zhang et al.~(2024) categorise 12,837 errors from 14 LLMs on
HumanEval, then replicate the dominant-category hierarchy on 23 models
across HumanEval Pro and MBPP Pro. The RFMDataset (RFM authors, 2025)
explicitly tests cross-model invariance and reports ``strikingly similar
failure mode distributions'' across 10 open-source models.

Beyond general-purpose evaluations, domain-specific taxonomies exist for
multi-hop QA (the Weakest Link Law authors (2026) show that multi-hop
performance collapses to the least-visible-evidence bucket on MuSiQue
and NeoQA), for multi-agent systems (MAST authors, 2025), and for
tool-augmented agents (\(\tau\)-bench authors, 2024). The convergent
finding across all these is the same: at any given corpus size, the
number of distinct named modes is small (typically 8--20), and the
cumulative coverage of the top modes is high. This is the empirical
phenomenon our theoretical framework aims to model.

\subsubsection{2.2 Targeted interventions}\label{targeted-interventions}

Each named failure mode in these taxonomies has, in some cases
independently, been the subject of focused intervention research. The
arithmetic-error cluster is the canonical example: PAL (Gao et al.,
2022) shows that delegating execution to a Python interpreter raises
GSM-Hard accuracy from 20.1\% (chain-of-thought) to 61.5\% (+41.4 pp),
with PoT (Chen et al., 2022) and ToRA (Gou et al., 2023) extending the
technique to broader math and financial reasoning. Constrained decoding
has effectively eliminated the format-violation cluster: DINGO (Suresh
et al., 2025) reports up to +68 pp on JSON-generation benchmarks, and
SLOT (SLOT authors, 2025) achieves 99.5\% schema accuracy while
preserving 94\% content similarity. Execution feedback closes most of
the code-error cluster: Reflexion (Shinn et al., 2023) improves
HumanEval pass@1 from 80\% to 91\%, and AgentCoder (Huang et al., 2023)
pushes it to 96.3\%. Process reward models substantially reduce
step-level reasoning errors: Math-Shepherd (Wang P. et al., 2023) raises
MATH performance from 28.6\% to 43.5\%, and Let's Verify Step by Step
(Lightman et al., 2023) reports 78\% on the MATH test set under process
supervision. Retrieval-augmented generation is the most studied of all:
Acurai (Acurai authors, 2024) reports 100\% hallucination elimination on
RAGTruth, with corroborating clinical and legal results elsewhere.
Preference optimisation closes the over-refusal gap: POROver (Karaman et
al., 2024) lifts the not-overrefused rate from 57.6\% to 82.1\% while
preserving 97.9\% safety. Structured uncertainty fixes most tool-call
failures: SAGE-Agent (Suri et al., 2025) raises When2Call accuracy from
36.5\% to 65.2\%.

Two recurring observations from this literature inform our theoretical
framework. First, each intervention is \emph{cluster-specific}: the
residual errors after intervention belong to a different named cluster,
not to the targeted one. Reflexion residuals are
specification-misinterpretation errors, not execution errors;
constrained-decoding residuals are semantic errors within valid
structure, not format violations. This is the empirical signature of the
cluster-orthogonality property that anchors Section 4.2. Second,
\emph{additivity is approximate but not perfect}. The Six Sigma Agent
framework reports compound reliability gains of \(14{,}700\times\) when
interventions target orthogonal failure modes (Six Sigma authors, 2026),
but Le (2026) shows that schema-level and prompt-level instructions
interact non-additively when they share a prompt channel. We handle this
caveat directly in Section 5.

\subsubsection{2.3 Length-decay benchmarks and their
interpretation}\label{length-decay-benchmarks-and-their-interpretation}

A parallel literature measures the \emph{shape} of the reliability decay
curve as context length grows. Two camps exist. The first reports decay
that is much milder than exponential. Loong (Wang M. et al., 2024) finds
GPT-4o on Chain-of-Reasoning tasks declines from 81.6\% (10--50K tokens)
to 32.9\% (\textgreater200K) --- a relative drop of about 60\% across
two doublings, consistent with a log-linear fit of accuracy
\(\approx 95 - 22\log_2(L/10\text{K})\). GSM-\(\infty\) (Zhou et al.,
2025) fits a sigmoid (saturating) decay with \(R^2 > 0.98\) and shows
that exponentially more inference compute yields only linear gains in
AUC --- a direct signature of clustered, not independent, errors. RULER
(Hsieh et al., 2024) finds threshold-like behaviour where simple
retrieval holds to 32K while complex reasoning collapses earlier,
inconsistent with smooth \((1-\varepsilon)^n\). Anchor-based LLMs (Pang
et al., 2024) achieve 99\% KV-cache reduction with only ``minor accuracy
compromise,'' directly implying that the effective decision count is
bounded.

The second camp reports ``rapid decay'' findings that are sometimes
cited as evidence for exponential compounding. Faith and Fate (Dziri et
al., 2023) shows multi-digit multiplication accuracy collapsing from
59\% at 3-digit to 4\% at 4-digit (zero-shot GPT-4) and argues
theoretically that ``the probability of incorrect predictions converges
exponentially to \(\approx 1\) for abstract compositional tasks.''
BABILong (Kuratov et al., 2024) reports that ``popular LLMs effectively
utilise only 10--20\% of the context and performance declines sharply
with increased reasoning complexity.'' METR (METR, 2025) reports that
the 80\%-success horizon is 4--6× shorter than the 50\%-success horizon,
suggesting a steep within-model reliability cliff. The Fano-style upper
bound of Wan et al.~(2025) explicitly theorises super-linear
information-demand growth and identifies an ``accuracy cliff'' at
capacity overflow.

The apparent tension between these two camps dissolves on careful
reading: every paper in the second group decays over a variable distinct
from raw token length \(n\). We document this in detail in Section 6.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Theoretical Framework}\label{theoretical-framework}

\subsubsection{\texorpdfstring{3.0 What \(|C|\) refers to: four
levels}{3.0 What \textbar C\textbar{} refers to: four levels}}\label{what-c-refers-to-four-levels}

LLM error analysis routinely conflates four distinct objects. To keep
the rest of this section honest, we distinguish them and tag the formal
claims accordingly.

\begin{itemize}
\tightlist
\item
  \textbf{L1: failure events} --- concrete LLM errors observed in a
  benchmark. These are the raw data behind any taxonomy.
\item
  \textbf{L2: empirical taxonomy categories} --- researcher-chosen
  labels grouping L1 events. ErrorAtlas's 17 categories, MWPES's top-4
  classes, and HumanEval's \texttt{AssertionError} / \texttt{NameError}
  partition belong here. L2 is observable but depends on the taxonomer's
  resolution choices.
\item
  \textbf{L3: latent failure modes} --- the (unobserved) underlying
  clusters that L2 categories approximate. L3 is the quantity Postulate
  1 is morally about, but it cannot be measured directly; we treat L2 as
  a noisy proxy for L3.
\item
  \textbf{L4: capability axes / interventions} --- the engineering unit.
  A Python interpreter, a constrained decoder, a retrieval-augmented
  generator. L4 is coarser than L2 in practice: one capability axis
  typically targets several L2 categories at once (Section 4.2.0).
\end{itemize}

Throughout this section, \(|C|\) refers to the \textbf{L2} count --- the
number of named taxonomy categories at a given corpus scale --- because
L2 is what published taxonomies report. Postulate 1 (Section 3.2) is a
statement about how L2 grows with sampled failures. Proposition 1
(Section 3.4) is engineering math conditional on Postulate 1. The
framework is \emph{engineering}-relevant because L4 is coarser than L2
(Section 3.2 closing); it is \emph{epistemically} conditional because L2
is a proxy for the actual L3 mode catalogue, and the proxy's
faithfulness has not been independently measured.

\subsubsection{\texorpdfstring{3.1 \(\beta\)-stratification of key
tokens}{3.1 \textbackslash beta-stratification of key tokens}}\label{beta-stratification-of-key-tokens}

The two-rate model of Arbuzov et al.~(2025), reproduced as Equation (1),
distinguishes \(k\) key tokens (error rate \(e_{\text{key}}\)) from
\(n - k\) non-key tokens (error rate
\(e_{\text{non}} \ll e_{\text{key}}\)). The empirical failure-mode
atlases, however, suggest that even within the key-token class errors
are not uniformly distributed. Most key tokens correspond to
``decisions'' for which the model has stable representations and
produces correct outputs; only a fraction concentrate the actual
failures.

\textbf{Definition 1 (Hard fraction).} Partition the \(k\) key tokens of
a sequence into easy and hard subsets:

\[
k_{\text{hard}} = \beta \cdot k, \qquad k_{\text{easy}} = (1 - \beta) \cdot k, \qquad \beta \in (0,1),
\]

where hard key tokens have an elevated error rate \(e_{\text{hard}}\)
corresponding to manifold-transition decisions (in the sense of Arbuzov
et al., 2025, §3.2), and easy key tokens have
\(e_{\text{easy}} \approx e_{\text{non}}\).

The composed reliability becomes

\[
P(\text{correct}) = (1 - e_{\text{hard}})^{\beta k} \cdot (1 - e_{\text{easy}})^{(1 - \beta) k} \cdot (1 - e_{\text{non}})^{n - k}.
\]

Because \(e_{\text{easy}} \approx e_{\text{non}}\) once context
accumulates, the failure rate is dominated by the \(\beta k\) hard
tokens, and the rest of this paper treats \(e_{\text{hard}}\) as the
load-bearing quantity.

\textbf{On the identifiability of \(\beta\).} \(\beta\) is a
\emph{latent} stratification parameter: it asks what fraction of
key-token \emph{decisions} are hard, not what fraction of \emph{observed
errors} fall into the most common named categories. Failure-mode atlases
report \(\Pr(\text{category} \mid \text{error occurred})\), which is a
different conditional from
\(\Pr(\text{hard} \mid \text{key-token decision})\). The ErrorAtlas
top-5-of-17 figure (53.6\%), the MWPES top-4 dominance (45--55\%), and
the HumanEval two-class coverage (86\%) are all consistent with
\(\beta\) being non-trivial --- they show failures are concentrated when
failures occur --- but they do not pin down \(\beta\). A more
identifying measurement would require ablating model confidence at
predicted decision boundaries, which the literature has not done at
scale. We therefore carry \(\beta\) symbolically through the rest of the
framework and report all downstream conclusions parametrically.

\subsubsection{3.2 Empirical postulate: logarithmic mode
discovery}\label{empirical-postulate-logarithmic-mode-discovery}

The \(\beta k\) hard-token failures across a corpus are not arbitrary:
they cluster into a finite set of recurring failure patterns indexed
\(c = 1, \ldots, |C|\). By the stratified-manifold argument of Arbuzov
et al.~(2025, §3.2), each pattern corresponds to a specific
manifold-transition mode, so the catalogue of modes is bounded in any
fixed domain.

The natural question is how \(|C|\) grows with the number of observed
failures. Two competing predictions exist in the type--token literature:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Heaps' law (standard).}
  \(|C|(k_{\text{hard}}) \approx K \cdot k_{\text{hard}}^{b}\) with
  typical exponent \(b \in [0.4, 0.6]\) (Manning et al., 2008). This is
  the canonical type--token growth law and is asymptotically compatible
  with Zipfian rank-frequency distributions.
\item
  \textbf{Logarithmic discovery.}
  \(|C|(k_{\text{hard}}) \leq A + \sigma \cdot \ln(k_{\text{hard}})\).
\end{enumerate}

\textbf{Important: Zipfian rank-frequency does not imply logarithmic
growth.} Although it is tempting to derive logarithmic mode discovery
from a Zipfian pattern-frequency distribution, the derivation fails:
Heaps' law is the correct type--token consequence of Zipf-distributed
events, and it is power-law, not logarithmic. We therefore state
logarithmic mode discovery as an empirical postulate, defensible by
direct measurement of error taxonomies, rather than as a theorem.

\textbf{Notation: corpus vs sequence.} Two sample-size variables matter
here, and conflating them is the seam that lets reviewers in. Let
\(h(n) = \beta k(n)\) denote the number of hard-token decision
opportunities in a single sequence of length \(n\). Let \(T\) denote the
number of observed hard-failure events in a corpus used to discover or
estimate the domain's failure catalogue. These are not the same object:
\(h(n)\) controls per-sequence exposure, while \(T\) controls empirical
catalogue discovery. We further write \(C_D\) for the full reachable
catalogue of recurring failure modes inside domain patch \(D\) (defined
at the end of this section), \(C_{\text{seen},D}(T)\) for the subset
discovered after \(T\) sampled failures, and \(C_{\text{active},D}(n)\)
for the subset that a single sequence of length \(n\) can activate.

\textbf{Postulate 1 (Patch-indexed catalogue discovery).} \emph{Within a
fixed application domain \(D\), the number of named recurring failure
modes discovered after \(T\) sampled hard-failure events is bounded by}

\[
|C_{\text{seen},D}(T)| \leq \min\bigl(A_D + \sigma_D \ln T,\ |C_D|\bigr), \qquad \sigma_D > 0,\ A_D \geq 0.
\]

\emph{Here \(T\) is a corpus-level sampling variable. The cap \(|C_D|\)
is the domain-imposed ceiling: the full reachable catalogue of recurring
modes inside patch \(D\). This postulate concerns catalogue discovery,
not the number of hard decisions inside a single sequence.}

\textbf{Assumption 2 (Per-sequence active-mode exposure).} \emph{For a
single sequence of length \(n\) in domain \(D\), let
\(C_{\text{active},D}(n)\) be the set of recurring failure modes that
can be activated by its \(h(n) = \beta k(n)\) hard decisions. We assume}

\[
|C_{\text{active},D}(n)| \leq \min\bigl(A'_D + \sigma'_D \ln h(n),\ |C_D|\bigr).
\]

\emph{This is a weaker, per-sequence analogue of Postulate 1. Longer
sequences expose the model to more active failure modes only slowly,
until the domain ceiling \(|C_D|\) is reached. The primed constants
\(A'_D\), \(\sigma'_D\) are distinct from those of Postulate 1 ---
corpus discovery and per-sequence activation need not share the same
rate. Proposition 1's sequence-length scaling uses
\(C_{\text{active},D}(n)\); full domain-library budgeting uses \(C_D\).}

\textbf{Empirical defense.} We anchor Postulate 1 against three
large-scale taxonomies:

\begin{itemize}
\tightlist
\item
  ErrorAtlas (Ashury-Tahan et al., 2026): 83 models × 35 datasets,
  \(k_{\text{hard}} \gtrsim 10^4\) failures, \(|C| = 17\). Taking
  \(A = 0\) as a conservative anchor,
  \(\sigma \approx 17 / \ln(10^4) \approx 1.85\).
\item
  HumanEval (Zhang et al., 2024): \(k_{\text{hard}} \approx 10^4\),
  \(|C| \in [8, 12]\) named modes, \(\sigma \in [1.0, 1.3]\).
\item
  MWPES (Sun et al., 2025): \(k_{\text{hard}} \approx 3 \times 10^5\),
  \(|C| \in [15, 20]\), \(\sigma \in [1.2, 1.6]\).
\end{itemize}

We adopt \(\sigma \approx 2\) as a target value with empirical range
\(\sigma \in [1.5, 3.0]\).

\textbf{Capabilities are coarser than error classes.} The σ estimate
above is derived from \emph{named error categories} in the published
taxonomies, but the capability harvest of Section 4.2 reveals that a
single capability often addresses multiple categories at once. A single
Python interpreter removes the execution-error component of arithmetic,
unit conversion, simple counting, list manipulation, and date
arithmetic, while leaving residual failures in problem representation
and code synthesis. A single constrained decoder eliminates by
construction the format-violation cluster and the structural component
of ``missing required element'' (ErrorAtlas's largest category at
15.56\%), because schemas with required fields cannot omit those fields.
Postulate 1's \(\sigma \approx 1.85\), calibrated against the raw
17-category taxonomy, is therefore conservative for the engineering
question: when interventions are coarse-grained, the practical
capability library is smaller than \(|C|\). The polylog conclusion of
Proposition 1 holds \emph{a fortiori}.

\textbf{Robustness.} The qualitative claim that follows in Section 3.4
is robust to the choice between Heaps' law (power-law) and logarithmic
discovery: as long as \(|C|\) grows sublinearly in \(k_{\text{hard}}\)
with an exponent less than one (or strictly slower), the resulting
intervention-budget bound is polylogarithmic in \(n\). We discuss the
Heaps variant explicitly in Section 3.5.

\textbf{Domain patches cap the available catalogue.} Postulate 1
describes how the \emph{count} of named modes grows with sampled
failures, but it leaves open whether the catalogue is unbounded
asymptotically. Real model deployment never visits the entire stratified
manifold that token representations occupy (Arbuzov et al.~2025, §3.2;
Li \& Sarwate 2025). A model deployed in a fixed application domain
\(D\) --- cardiology RAG, legal contract drafting, code review ---
occupies a bounded sub-region of that manifold, which we call the
\emph{domain patch} \(P_D\). The patch may contain arbitrarily complex
local transition geometry; it does not require the local complexity
inside the patch to be small. What the patch does is restrict the
\emph{support} of reachable failure modes. A cardiology RAG agent will
not produce poetry-metaphor failures or theorem-proving failures; its
accessible catalogue is bounded by what its operational neighborhood
actually admits.

Let \(C_D\) denote the set of recurring failure modes accessible within
\(P_D\). Then empirical mode discovery is bounded by

\[
|C_{\text{observed}}(t)| \leq \min\bigl(A_D + \sigma_D \ln t,\ |C_D|\bigr).
\]

The logarithmic term is Postulate 1 applied within the patch; the cap
\(|C_D|\) is the domain-imposed ceiling. We do \emph{not} claim that the
patch derives logarithmic discovery --- the patch bounds the catalogue
and makes logarithmic or saturating discovery dynamics plausible inside
fixed domains. The cap is what changes the asymptotic behaviour of
Proposition 1: once enough failures have been sampled that
\(A_D + \sigma_D \ln k_{\text{hard}}\) reaches \(|C_D|\), further
sampling does not enlarge the catalogue and the intervention budget
becomes a domain-constant function of \(|C_D|\) alone.

Two lines of evidence support the cap. Structurally, embedding-space
analyses find that representations decompose into local strata whose
intrinsic dimensions range \(\sim 5\)--\(15\) across domains, well below
the ambient \(\sim 10^3\) scale (Park et al.~2024; Li \& Sarwate 2025)
--- the differential-geometric anchor for treating \(P_D\) as a
sub-manifold rather than a slogan. Empirically, cross-domain
heterogeneity is now routine: MMLU-Pro reports a 14-domain accuracy
spread from \(\sim 20\%\) to \(> 70\%\) on a single base model (Wang Y.
et al.~2024); Kandpal et al.~(2023) find log-linear scaling between
training-document count and retrieval accuracy
(\(R^2 \in [0.98, 0.99]\)); Mallen et al.~(PopQA, 2023) identify a
log-popularity threshold below which scaling does not help. None of
these \emph{measures} \(|C_D|\) directly. Together they motivate the
patch-indexed hypothesis without measuring \(C_D\); direct support would
require per-domain mode-discovery curves. The cap is currently a
structural conjecture consistent with the observed heterogeneity, not an
empirical fit.

Crucially, \(\sigma_D\), \(A_D\), \(\beta_D\), and \(|C_D|\) are all
domain-indexed. Deploying the same base model in three different
application domains yields three distinct mode catalogues --- not
because the model is different but because the patches are. Reliability
engineering is therefore not global manifold mastery; it is \emph{local
patch coverage}: identify the recurring transition types available in
the target patch and provision interventions against the head of that
catalogue. We return to this engineering implication in Section 5 and to
the patch-shift caveat in Section 6.3.

The discovery variable \(T\), the sequence-exposure variable \(h(n)\),
and the patch ceiling \(|C_D|\) answer different questions: \(T\)
determines how much of the catalogue we have measured, \(h(n)\)
determines how much of it a single sequence can activate, and \(|C_D|\)
determines the maximum catalogue that the domain can expose.

\subsubsection{3.3 Coverage by a targeted intervention
library}\label{coverage-by-a-targeted-intervention-library}

Given a catalogue of \(|C|\) recurring modes, an \emph{intervention
library} of size \(m\) covers the \(m\) most prevalent modes, leaving
the remaining \(|C| - m\) uncovered. The cumulative coverage of such a
library depends on the rank-frequency distribution of the patterns
themselves.

For Zipf-distributed pattern frequencies with exponent \(s = 1\), the
discrete cumulative mass is \(H_m / H_{|C|}\) where
\(H_j = \sum_{i=1}^{j} 1/i\) is the harmonic number. For our ErrorAtlas
anchor (\(|C| = 17\), \(m = 5\)), this gives
\(H_5 / H_{17} \approx 66.4\%\), which overestimates the observed
coverage of 53.6\%.

A closer empirical fit to the ErrorAtlas data is

\[
F(m; |C|) = \min\!\left(1,\ \frac{\ln m}{\ln |C|}\right),
\]

which yields \(F(5; 17) \approx 56.8\%\) and
\(F(10; 17) \approx 81.3\%\), matching the observed 53.6\% and 80\%+
within three percentage points. This is a convenient empirical-fit form
among several plausible candidates --- not a derived consequence of Zipf
and not a privileged choice on theoretical grounds. Alternative
cumulatives (Zipf--Mandelbrot with offset, truncated power law,
saturating exponential) can be fit to the same two anchor points with
comparable error, and a structural preference among them is not
available at current evidence levels. We adopt the \(\ln(m)/\ln(|C|)\)
form for the rest of the paper because it is the simplest functional
form that respects the \(\min(1, \cdot)\) cap and reproduces the
ErrorAtlas anchors. Section 3.6 shows that the qualitative polylog
conclusion of Proposition 1 survives across all four candidate
cumulatives we consider.

A second caveat is that \(F(m; |C|)\) counts \emph{named L2 categories
covered}, not \emph{L4 capability axes deployed}. A single capability
axis typically targets several L2 categories on average (Section 4.2.0).
Real-deployment residual error is therefore often lower than
\(F(m; |C|)\) predicts at any given \(m\): the bound remains valid but
is loose in this regime. Empirically observed library sizes that achieve
a given residual threshold may be a constant factor smaller than the
formula's prediction.

After deploying the top-\(m\) intervention library, the residual
per-hard-token error rate is

\[
e_{\text{res}}(m) = \bigl(1 - F(m; |C|)\bigr) \cdot e_{\text{hard}}.
\]

\subsubsection{3.4 Main theorem: polylogarithmic intervention
budget}\label{main-theorem-polylogarithmic-intervention-budget}

We now compose the three ingredients --- \(\beta\)-stratification, the
patch framework of Section 3.2, and the coverage form --- with the
two-rate model from Arbuzov et al.~(2025).

\textbf{Asymptotic regime.} For sequence-length scaling we use
Assumption 2 (per-sequence active-mode exposure), not Postulate 1
(corpus-level catalogue discovery). The empirical sparsity anchor
\(\alpha = k/n \in [0.05, 0.10]\) is a \emph{finite-length} statement at
typical \(n\); under Arbuzov et al.~(2025, §3.1)'s \(k \sim \log n\)
regime, \(\alpha\) decays as \(O(\log n / n)\). Substituting
\(h(n) = \beta k(n)\) into the active-mode bound of Assumption 2 gives

\[
|C_{\text{active},D}(n)| \leq \min\bigl(A'_D + \sigma'_D \ln(\beta k(n)),\ |C_D|\bigr).
\]

If \(k(n) = \Theta(\log n)\), then before the patch cap is reached,
\(|C_{\text{active},D}(n)| = O(\log\log n)\). Once the cap is reached,
\(|C_{\text{active},D}(n)| = |C_D|\) and the required library size
becomes a domain-constant independent of \(n\). If \(k \leq k_{\max}\)
(the bounded regime of Arbuzov et al., 2025, §3.1),
\(|C_{\text{active},D}|\) is constant in \(n\) throughout.

\textbf{Proposition 1 (Per-Hard-Token Intervention Budget).} \emph{Let
\(\varepsilon \in (0, e_{\text{hard}})\) be a target per-hard-token
residual error rate for sequences in domain \(D\). Let
\(C_{\text{eff}}\) denote the relevant covered catalogue:}

\begin{itemize}
\tightlist
\item
  \emph{\(C_{\text{eff}} = C_{\text{active},D}(n)\) for per-sequence
  reliability scaling, or}
\item
  \emph{\(C_{\text{eff}} = C_D\) for full domain-library deployment.}
\end{itemize}

\emph{Under the coverage form of Section 3.3, the smallest intervention
library satisfying \(e_{\text{res}}(m) \leq \varepsilon\) has size}

\[
m \geq |C_{\text{eff}}|^{1 - \varepsilon / e_{\text{hard}}}.
\]

\emph{If \(C_{\text{eff}} = C_{\text{active},D}(n)\) and
\(k(n) = \Theta(\log n)\), then in the pre-cap regime}

\[
m = O\!\bigl((\log\log n)^{1 - \varepsilon / e_{\text{hard}}}\bigr),
\]

\emph{i.e., the intervention budget grows doubly logarithmically with
sequence length. In the cap regime, or for full domain-library
deployment, \(m \geq |C_D|^{1 - \varepsilon / e_{\text{hard}}}\),
independent of \(n\).}

\emph{Proof.} From Section 3.3,
\(e_{\text{res}}(m) = (1 - F(m; |C_{\text{eff}}|))\,e_{\text{hard}}\).
Requiring \(e_{\text{res}}(m) \leq \varepsilon\) gives
\(F(m; |C_{\text{eff}}|) \geq 1 - \varepsilon / e_{\text{hard}}\). Since
\(\varepsilon < e_{\text{hard}}\), the required coverage lies below one,
so the cap in \(F\) is inactive. With
\(F(m; |C_{\text{eff}}|) = \ln m / \ln |C_{\text{eff}}|\), we obtain
\(\ln m / \ln |C_{\text{eff}}| \geq 1 - \varepsilon / e_{\text{hard}}\),
hence \(m \geq |C_{\text{eff}}|^{1 - \varepsilon / e_{\text{hard}}}\).
For per-sequence scaling, Assumption 2 gives
\(|C_{\text{eff}}| = |C_{\text{active},D}(n)| = O(\log\log n)\) when
\(k(n) = \Theta(\log n)\), before the patch cap is reached. After the
cap is reached, \(|C_{\text{eff}}| = |C_D|\) and the bound is
independent of \(n\). \(\square\)

We label this a \emph{Proposition} rather than a \emph{Theorem} to keep
its conditional nature visible. The formal content is engineering math,
derived from the empirical postulates of Sections 3.2 and 3.3; it does
not derive a law of LLM reliability so much as formalise a directional
intuition. Section 3.5 already records the standard Heaps-power-law
alternative; Section 3.6 below shows the qualitative polylog conclusion
survives across a wider family of cluster-count laws. The
doubly-logarithmic rate stated above is the \emph{optimistic special
case} --- the per-sequence pre-cap regime under logarithmic mode
discovery; weaker (but still polylog) rates obtain under Heaps; cap
regime is a domain-constant.

\textbf{Sequence-level versus per-hard-token bounds.} Proposition 1
bounds the \emph{per-hard-token} residual error. Sequence-level failure
probability is a strictly stronger target. Taking logs of the composed
reliability and using \(\log(1 - x) \approx -x\) for small \(x\),

\[
-\log P(\text{correct}) \approx \beta k \cdot e_{\text{res}}(m) + (1 - \beta) k \cdot e_{\text{easy}} + (n - k) \cdot e_{\text{non}}.
\]

The first term dominates \textbf{only under the assumption that
\(e_{\text{non}} = o(k/n)\) asymptotically} --- i.e., that non-key
tokens have vanishing per-token error rate fast enough that their
aggregate \((n-k) \cdot e_{\text{non}}\) stays below
\(\beta k \cdot e_{\text{res}}(m)\). Without this assumption, the
non-key term can re-introduce the original exponential-in-\(n\) concern
through the back door. We adopt this assumption --- it is the natural
reading of Arbuzov et al.~(2025)'s observation that non-key tokens
become ``nearly deterministic'' once context accumulates --- but we name
it explicitly because the literature has not measured
\(e_{\text{non}}(n)\) directly. Where this assumption is unsafe (e.g.,
domains with persistent baseline noise on routine tokens), Proposition 1
should be read as a bound on \emph{key-token failure risk} only, not on
literal sequence-level correctness.

Granted that assumption, bounding sequence-level failure probability by
\(\varepsilon_{\text{seq}}\) requires
\(e_{\text{res}}(m) \leq \varepsilon_{\text{seq}} / (\beta k)\).
Substituting into the budget formula,

\[
m \geq |C|^{1 - \varepsilon_{\text{seq}} / (\beta k \cdot e_{\text{hard}})}.
\]

This is a substantially stronger requirement than Proposition 1: as
\(k\) grows, the per-token tolerance tightens and \(m\) approaches
\(|C|\) --- essentially every named mode must be covered. The
widely-quoted ``fix \(\approx 50\) patterns per domain →
\(\approx 90\%\) error reduction'' claim is best understood as a
per-hard-token statement; the analogous sequence-level claim requires
correspondingly more interventions and tighter coverage of the tail.
Production systems with one-shot reliability targets should plan for the
stricter regime.

\subsubsection{3.5 Alternative formulation: Heaps power-law
variant}\label{alternative-formulation-heaps-power-law-variant}

A reader who prefers to derive cluster-count growth from standard Heaps'
law rather than from Postulate 1 obtains a qualitatively similar result.
Let \(|C|(k_{\text{hard}}) \approx K \cdot k_{\text{hard}}^{b}\) with
\(b \in (0, 1)\). Canonical fits give \(b \approx 0.5\) for
natural-language vocabularies; failure-mode taxonomies plateau much more
sharply (ErrorAtlas stabilises at \(|C| = 17\) across \(10^4{+}\)
failures), implying a smaller effective \(b \approx 0.2\) in our
setting. Composing with \(k = \Theta(\log n)\), we get
\(|C| = O((\log n)^{b})\), and Proposition 1 becomes

\[
m = O\!\left((\log n)^{b \cdot (1 - \varepsilon / e_{\text{hard}})}\right).
\]

This is still polylogarithmic in \(n\) for any \(b \in (0, 1)\) and any
\(\varepsilon < e_{\text{hard}}\). The paper's qualitative claim ---
that the intervention budget grows polylogarithmically in sequence
length --- survives either choice of cluster-count law.

\subsubsection{3.6 Sensitivity to the cluster-count law: symbolic
form}\label{sensitivity-to-the-cluster-count-law-symbolic-form}

The polylog conclusion of Proposition 1 depends on which cluster-count
law one accepts. The available evidence (Section 3.2) is consistent with
multiple candidates --- no subsample-discovery curve has been published
for any LLM failure-mode taxonomy at the time of writing. We therefore
report symbolic rates rather than fitted constants. Assume
\(h(n) = \beta k(n)\) is the number of hard decisions activated by a
sequence. The candidate active-catalogue laws then imply:

\begin{itemize}
\tightlist
\item
  \textbf{Logarithmic:} \(|C_{\text{active},D}(n)| = O(\log h(n))\). If
  \(k(n) = \Theta(\log n)\), then
  \(m = O\!\bigl((\log\log n)^{1 - \varepsilon / e_{\text{hard}}}\bigr)\).
\item
  \textbf{Heaps:} \(|C_{\text{active},D}(n)| = O\!\bigl(h(n)^b\bigr)\)
  with \(b \in (0, 1)\). If \(k(n) = \Theta(\log n)\), then
  \(m = O\!\bigl((\log n)^{b\,(1 - \varepsilon / e_{\text{hard}})}\bigr)\).
\item
  \textbf{Saturating:} \(|C_{\text{active},D}(n)| \leq |C_D|\). Then
  \(m = O\!\bigl(|C_D|^{1 - \varepsilon / e_{\text{hard}}}\bigr)\),
  constant in \(n\) once the patch ceiling is reached.
\end{itemize}

The qualitative conclusion is robust: under every candidate law, \(m\)
grows more slowly than any positive power of \(n\), and the directional
claim (``a small library covers the head of the failure distribution in
the per-hard-token regime'') survives. Only the exponent shifts:
doubly-logarithmic under logarithmic discovery, \((\log n)^b\) with
small \(b\) under Heaps, constant in the cap regime. The
doubly-logarithmic rate is the optimistic special case.

Numerical constants require a measured discovery curve
\(C_{\text{seen},D}(T)\) or \(C_{\text{active},D}(n)\). Existing
taxonomies provide endpoint category counts at a single corpus scale
(e.g., ErrorAtlas at \(|C| = 17\) for \(\approx 10^4\) failures), not
discovery curves. We therefore report rates rather than fitted
constants, and treat the subsample-discovery curve as the explicit
empirical test that would either tighten the postulate or fall back to
the Heaps variant. Until that measurement exists, the framework's
headline rate should be read as \emph{polylogarithmic in the pre-cap
regime, domain-constant in the cap regime} --- with the specific
exponent flagged as a falsifiability test rather than a fitted
prediction.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Empirical Evidence}\label{empirical-evidence}

We now collect quantitative evidence for the three load-bearing claims:
errors cluster into a small recurring set (Section 4.1); each cluster is
independently addressable by one targeted intervention (Section 4.2);
and reliability decays sublinearly, not exponentially, with output
length (Section 4.3).

\subsubsection{4.1 Claim A: Failure-mode
clustering}\label{claim-a-failure-mode-clustering}

\textbf{Cross-domain concentration.} The strongest single result is
ErrorAtlas (Ashury-Tahan et al., 2026). Across 83 models and 35
datasets, with approximately 7,000 sampled failures, just five of the
seventeen named categories account for 53.6\% of all errors, and ten of
seventeen for over 80\%. The category set is fixed in advance; the
cumulative coverage curve is what models and datasets produce when
binned into it.

\textbf{Domain-specific Paretos.} The cross-domain picture is reproduced
at finer grain within each domain:

\begin{itemize}
\tightlist
\item
  \textbf{Math.} MWPES-300K (Sun et al., 2025) categorises 304,865
  errors from 15 LLMs across four math word-problem datasets. Per-model,
  the top four categories capture 45--55\% of errors. Critically, the
  \emph{same} top-four set appears across all 15 models --- a
  dataset-invariance result.
\item
  \textbf{Code.} Zhang et al.~(2024) categorise 12,837 errors from 14
  LLMs on HumanEval. \texttt{AssertionError} and \texttt{NameError}
  alone capture 86.35\% of failures. The dominant-category hierarchy
  replicates across 23 models on HumanEval Pro and MBPP Pro.
\item
  \textbf{Multi-hop QA.} The Weakest Link Law (Weakest Link authors,
  2026) shows multi-hop performance on MuSiQue and NeoQA collapsing to
  the least-visible-evidence bucket --- a single dominant failure mode
  (missing evidence retrieval) accounting for most errors across
  question types.
\item
  \textbf{Agentic / tool use.} MAST (MAST authors, 2025) and
  \(\tau\)-bench (TauBench authors, 2024) document recurring failure
  modes in tool-augmented agents (tool selection, parameter
  disambiguation, intermediate state loss). The mode count is small
  (under twenty in each study) and the cumulative coverage of the top
  modes is high.
\item
  \textbf{Retrieval-augmented generation.} Acurai (Acurai authors, 2024)
  on RAGTruth and follow-on clinical/legal RAG work all report that
  hallucinations cluster into a handful of distinguishable modes
  (citation hallucination, context-stitching errors, retrieval-context
  mismatch), each independently addressable.
\end{itemize}

\textbf{Cross-model invariance.} The same top categories appear across
very different models. ErrorAtlas reports a stable category hierarchy
across 83 models. RFMDataset (RFM authors, 2025) explicitly tests
invariance and reports ``strikingly similar failure mode distributions''
across 10 open-source models. EDIT (Dai et al., 2025) finds an explicit
\(\approx 4.7\%\) key-step fraction across reasoning chains and across
models, with stable category labels.

\textbf{Theoretical support.} Schaeffer et al.~(2025) prove that the
observed aggregate power-law scaling of LLM evaluations requires the
per-problem success-rate distribution to be heavy-tailed near \(p = 0\).
This is not a direct theorem about cluster-count growth (which is the
object of our Postulate 1), but it is a closely related structural
result: the failure side of LLM behaviour is intrinsically heavy-tailed.

The aggregate picture is what Postulate 1 of Section 3.2 is designed to
capture: LLM errors are not a long uniform tail but a short, repeating
list. The ErrorAtlas \(|C| = 17\) with 80\% top-10 coverage is the
cleanest summary; per-domain Paretos show the same shape inside each
domain; the cross-model invariance shows the catalogue is stable.

\subsubsection{4.2 Claim B: Cluster-orthogonal
interventions}\label{claim-b-cluster-orthogonal-interventions}

The intervention literature provides at least one targeted
countermeasure for each named cluster of Section 4.1, almost always with
a published quantitative result. Two organising granularities exist in
this evidence, and we report both. At the \emph{capability} level (six
axes; Section 4.2.0) the cluster-orthogonality property is most clearly
visible --- capability provisioning leaves a class-shift signature that
says exactly which class was erased and which classes the residuals
belong to. At the \emph{error-class} level (twelve named clusters;
Sections 4.2.1--4.2.2) the evidence aligns with the taxonomies of
Section 4.1 but is finer-grained than the underlying capability
mechanisms. The two views are complementary: the axes are the
engineering unit, the clusters are the bookkeeping unit.

\paragraph{4.2.0 Six capability axes and three structural
patterns}\label{six-capability-axes-and-three-structural-patterns}

A dedicated harvest of capability-elimination evidence (parallel to the
cluster-based harvest below) yields \textbf{27 GOLD-tier citations}
across \textbf{six independent capability axes}: (i) Arithmetic
(Python/symbolic execution), (ii) Code Execution (REPL/sandbox
feedback), (iii) Format/Structure (constrained decoding, FSMs, grammar
engines), (iv) Perception/Grounding (visual grounding models for GUI,
charts, tables), (v) Knowledge/RAG (inference-time dense retrieval and
citation grounding), and (vi) Verification (formal proof checkers,
learned verifiers, classifier rerouting). Each axis is independently
confirmed by 3--9 GOLD citations. The 27 citations exhibit three
structural patterns of capability-elimination that strengthen --- and
refine --- the cluster-orthogonality claim.

We stratify these into three tiers by the \emph{kind} of evidence each
provides, taking care not to inflate empirical reductions into
mathematical guarantees.

\textbf{Pattern A --- Hard guarantees (by-construction).} Six citations
achieve residual error rate equal to zero \emph{by construction},
restricted strictly to structural / verifiable error classes:
constrained decoders set \(P(\text{invalid token}) = 0\) at every
generation step (DINGO, ToolDec, XGrammar, OpenAI Structured Outputs);
proof kernels reject any output that fails type-checking (Lean /
DeepSeek-Prover-V2). The class of ``outputs that violate the grammar''
is mathematically empty under these mechanisms. We restrict Pattern A to
this narrow sense: format, syntax, schema, formal-proof validity.

\textbf{Pattern B --- Strong empirical reductions with class-shift
signature.} Roughly fourteen citations report empirical reductions of
80--100\% in a named error class, with the post-intervention failure log
dominated by structurally different residual classes. The cleanest
example is Program-of-Thoughts on GSM8K (Chen et al., 2022): manual
taxonomy of 50 PoT failures shows calculation errors drop from 30\% of
CoT failures to \textbf{0\%}, while residuals are 62\% reasoning + 36\%
misunderstanding. In the OpenMedCalc arm of Goodell et al.~(2025),
\emph{``only interpretation errors were identified''} --- the
calculation-error class is empirically absent from the post-tool failure
log. Acurai (2024) reports 100\% hallucination elimination on RAGTruth
(95\% CI 91--100\%) for GPT-4 and GPT-3.5 --- a strong empirical result
on that benchmark, not a by-construction property of RAG in general. The
class-shift signature is the structural observation: post-intervention
residuals belong to different classes, consistent with cluster
orthogonality, though formally the framework only requires this in
expectation rather than identically.

\textbf{Pattern C --- Moderate reductions (60--80\%) with residuals
shifting outside the target class.} Several citations: Legal RAG
(Dantart, 2026) reduces fabricated citations from \textgreater30\% to
\textless0.2\%; SimpleQA web search (OpenAI, 2025) cuts factual errors
from 47\% to 9.6\%; CRITIC toxicity (Gou et al., 2024) reduces toxic
generation by 79.2\%. Residuals shift toward reasoning / planning errors
rather than inflating the targeted class. This is the regime where the
framework's ``selectivity'' claim is supported but does not rise to
elimination.

The tier boundaries matter for what the paper claims. Pattern A admits
the strongest rhetoric (``by construction'', ``mathematically empty'').
Pattern B admits ``strong empirical reduction'' and ``class-shift
signature'' but not ``by construction'' --- Acurai's 100\% on RAGTruth
is a benchmark-conditional empirical fact, not a mathematical guarantee.
Pattern C supports ``selective reduction'' without claiming class
elimination. The earlier framing in this section's first draft blurred
these boundaries; the revised version keeps them sharp.

The 27 GOLD citations, organised by axis and pattern, are summarised
below.

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Axis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Targeted class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Pre → Post
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Pattern
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Verification & DeepSeek-Prover-V2 (2025) & Invalid Lean proof & any →
0\% by construction & A \\
Format & OpenAI Structured Outputs (2024) & JSON schema violation &
\textgreater60\% → 0\% by construction & A \\
Format & ToolDec (Zhang et al., 2023) & Tool-call syntax error &
21--100\% → 0\% & A \\
Format & DINGO (Suresh et al., 2025) & JSON parse failure & 13--82\% →
0\% & A \\
Format & XGrammar (Dong et al., 2024) & Multi-format errors & 20--38\% →
0\% & A \\
Format & XGrammar-2 (Dong et al., 2026) & Malformed tool calls &
33--78\% → 0\% & A \\
Arithmetic & Program of Thoughts (Chen et al., 2022) & Calculation
errors on GSM8K & 30\% of failures → 0\% & B \\
Arithmetic & OpenMedCalc (Goodell et al., 2025) & Clinical arithmetic &
``only interpretation errors remained'' & B \\
Knowledge/RAG & Radiology RAG (Wada et al., 2025) & RAG hallucinations &
8\% → 0\% (p=0.012) & B \\
Knowledge/RAG & Acurai (2024) & Context-conflict hallucinations & 100\%
→ 0\% on conflict subset & B \\
Knowledge/RAG & Rigorous Archivist (Dantart, 2026) & Fabricated legal
citations & \textgreater30\% → \textless0.2\% & C \\
Code Exec & LlmFix (Zhang et al., 2024) & SyntaxError (HumanEval) &
5.76\% → 0.01\% & A \\
Code Exec & AlphaCode (Li et al., 2022) & False-positive submissions &
62\% → 4\% & B \\
Code Exec & MGDebugger (Shi et al., 2024) & 5 code-bug classes & 100\%
repair (5/6 classes) & B \\
Knowledge/RAG & ALCE (Gao et al., 2023) & Citation grounding errors &
73\% ungrounded → 15\% & C \\
Knowledge/RAG & GPT-5 SimpleQA (OpenAI, 2025) & Factual errors (web QA)
& 47\% → 9.6\% & C \\
Knowledge/RAG & Almanac (NEJM AI 2024) & Clinical citation errors & 44\%
→ 9\% & C \\
Arithmetic & MedRaC (Liu et al., 2025) & Arithmetic in medical reasoning
& 426 → 74 errors & B \\
Code Exec & LlmFix NameError (Zhang et al., 2024) & NameError repair &
22.7\% → 2.3\% & A \\
Perception & UGround (Gou et al., 2025) & GUI grounding errors & 16.2\%
→ 73.3\% accuracy & B \\
Perception & Jedi / OSWorld (Xie et al., 2025) & GUI task failures & 5\%
→ 27\% SR (5.4×) & B \\
Perception & SeeClick (Cheng et al., 2024) & Element mis-location &
5.2\% → 53.4\% (10.3×) & B \\
Perception & DePlot (Liu et al., 2023) & Chart-reading errors & 38.2\% →
67.6\% (+29.4 pp) & B \\
Verification & CRITIC toxicity (Gou et al., 2024) & Toxic generation &
79.2\% reduction & C \\
Knowledge/RAG & Self-RAG (Asai et al., 2024) & Unsupported facts &
55.5\% → 22\% error & B \\
Perception & OmniParser (Lu et al., 2024) & Icon / element errors &
70.5\% → 93.8\% (79\% error reduction) & B \\
Knowledge/RAG & PopQA (Mallen et al., 2023) & Long-tail entity errors &
\textasciitilde80\% → \textasciitilde50\% failure (\textasciitilde75\%
on tail) & B \\
\end{longtable}
}

\textbf{Capabilities are coarser than error classes.} Several GOLD
citations reveal ``two-for-one'' reductions where one capability
addresses multiple named clusters. A single Python interpreter removes
the execution-error component of arithmetic (cluster A), unit conversion
(B), simple counting (C), list manipulation, and date arithmetic ---
five named classes under one provisioning, with residual failures
shifting to problem representation and code synthesis. Constrained
decoding eliminates by construction both format violations (D) and the
structural component of ``missing required elements'' (the largest
single ErrorAtlas category at 15.56\%, our Category I), because schemas
with \texttt{required:\ {[}...{]}} fields cannot omit those fields. Code
execution feedback strongly reduces SyntaxError, NameError, and most
TypeError (three subclasses of cluster E) together. RAG strongly reduces
factual hallucinations (J), fabricated citations, and outdated
information jointly when the relevant evidence is retrieved and
grounded, while leaving residual failures in retrieval coverage, context
selection, and reasoning over retrieved material. The practical
capability library required to address a given fraction of failures is
therefore smaller than the count of named error categories --- the
empirical phenomenon noted in Section 3.2 as a conservative-bound
tightening of Postulate 1.

A negative-control result confirms the framework's selectivity:
DebugBench (Tian et al., 2024) finds that execution feedback works well
for syntax/reference errors but is \emph{``unhelpful for logic
errors\ldots{} may even cause disruptions.''} Capability provisioning
addresses what it can detect or constrain, and the prediction of where
it fails (semantic / open-ended classes) is correct. We return to this
point in Section 6.

\paragraph{Summary of the twelve
categories}\label{summary-of-the-twelve-categories}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Cluster
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Covering axis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Intervention
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Before → After
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\(\Delta\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Strength
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
A. Arithmetic / computation & Arithmetic & Python interpreter (PAL) &
20.1\% → 61.5\% & +41.4 pp & Strong & Gao et al., 2022 \\
B. Unit conversion & Arithmetic & Folded into A (Python handles units
natively) & --- & --- & (Folded) & --- \\
C. Counting / enumeration & Arithmetic & Explicit counter (research gap)
& --- & --- & Gap & Mechanistic interp. work \\
D. Format / schema violation & Format/Structure & Constrained decoding
(DINGO, SLOT) & baseline → +68 pp; 99.5\% schema acc. & +68 pp & Strong
& Suresh et al., 2025; SLOT authors, 2025 \\
E. Code logic / type / null & Code Exec & REPL feedback (Reflexion,
AgentCoder) & 80\% → 96.3\% pass@1 & +16.3 pp & Strong & Shinn et al.,
2023 \\
F. Multi-hop decomposition drift & Knowledge/RAG & Entity-grounded
rewriter (ChainRAG) & consistent gains across 3 benchmarks &
\$\sim\$5--10 pp & Medium & Zhu et al., 2025 \\
G. Reasoning step errors & Verification & Process reward model
(Math-Shepherd) & 28.6\% → 43.5\% & +14.9 pp & Strong & Wang P. et al.,
2023 \\
H. Specification misinterpretation & (semantic residual) & Clarification
loops (AmbigNLG) & ROUGE-L +15; compliance 86\% → 94\% & +8 pp & Medium
& Niwa \& Iso, 2024 \\
I. Missing required elements & Format/Structure & Reframed as Category D
for structural, G/H for semantic & (subsumed) & --- & Reframed & This
work \\
J. Hallucination / factual errors & Knowledge/RAG & RAG query
reformatting (Acurai) & non-zero → 0\% & \(-100\%\) & Strong & Acurai
authors, 2024 \\
K. Inappropriate refusal & Verification & Preference optimisation
(POROver) & 57.6\% → 82.1\% & +24.5 pp & Strong & Karaman et al.,
2024 \\
L. Tool / API usage errors & Format/Structure + Verification &
Structured uncertainty (SAGE-Agent) & 36.5\% → 65.2\% & +28.7 pp &
Strong & Suri et al., 2025 \\
\end{longtable}
}

Eight of twelve categories have a strong citation with double-digit
percentage-point improvement; three (B, C, I) lack a clean
single-cluster ablation in the literature and are addressed below; one
(F) has consistent medium-strength evidence.

\paragraph{Detailed evidence on the strongest
categories}\label{detailed-evidence-on-the-strongest-categories}

The arithmetic category is the cleanest example of cluster
orthogonality. Delegating execution to a Python interpreter, PAL (Gao et
al., 2022) lifts GSM-Hard accuracy from 20.1\% to 61.5\% --- a
41.4-point gain on the benchmark configuration where large numbers
stress LLM tokenisation. What matters more than the gain is the
composition of the residual: PAL's remaining failures are
problem-comprehension errors, mis-translations from word problem to
code. Arithmetic execution errors have left the failure log. Toolformer
(Schick et al., 2023) self-teaches the same shift and lands a 97.9\%
calculator-call rate on arithmetic queries; and on clinical calculations
Goodell et al.~(2025) report a 5--13× error reduction (LLaMA: 16\% →
88\%; GPT-4: 4.8\% → 64\%) once the appropriate domain tool is in the
loop.

Format violations admit something stronger still: a mathematical
guarantee. DINGO (Suresh et al., 2025) reports up to +68 percentage
points on JSON-generation tasks. SLOT (SLOT authors, 2025) reaches
99.5\% schema accuracy with 94\% content similarity --- the cluster is
essentially gone. A systematic evaluation of four production frameworks
(Guidance, Outlines, Llamacpp, XGrammar) by JSONSchemaBench (Geng et
al., 2025) reports coverage up to 0.96 on real-world schemas. The
mechanism is uniform across the family: zero out the logits of
grammar-violating tokens at each decoding step and the violations cannot
occur.

The hallucination result is the most striking single number in the
harvest. Acurai (Acurai authors, 2024) --- query reformatting applied to
RAG --- reports 100\% hallucination elimination on RAGTruth for both
GPT-4 and GPT-3.5, with a 95\% CI of 91--100\%. This is a
benchmark-specific empirical result, not a general by-construction
guarantee: Acurai does not prove that hallucinations cannot occur, only
that none were observed on RAGTruth's test set under their specific
query-reformatting setup. Domain-specific variants (legal, clinical)
report 60--90\% reduction with magnitudes depending on the baseline. The
mechanism is austere but not mathematical: hallucinations require the
model to generate factual claims absent from context, and
citation-grounded retrieval tightens --- without eliminating in general
--- that failure path.

Code generation shows a clean staircase. Reflexion (Shinn et al., 2023)
lifts HumanEval pass@1 from 80\% to 91\% by reflecting on execution
feedback; AgentCoder (Huang et al., 2023) extends the loop to
multi-agent test-driven generation and reaches 96.3\% --- approaching
ceiling. At 96.3\% the residual is dominated by spec-misinterpretation
(Category H). The execution-feedback capability has erased what it can
erase.

\paragraph{The Category-I reframing}\label{the-category-i-reframing}

Of the three apparent gaps (B, C, I), Category I --- ``missing required
elements'' --- is the most significant because it is the largest single
category in ErrorAtlas at 15.56\% of all errors. Originally we expected
to find a dedicated ``checklist verifier'' intervention in the
literature; we did not. The cleanest resolution is mechanistic
equivalence with Category D. A ``missing required element'' error is
structurally a schema-violation where the schema's
\texttt{required:\ {[}...{]}} field list is not satisfied. Constrained
decoders (DINGO, SLOT, Outlines, Guidance, XGrammar) enforce schema
constraints during decoding; with a properly specified schema, they
mathematically \emph{cannot} produce output that omits a required field.
SLOT's 99.5\% schema accuracy directly includes \texttt{required}
enforcement.

We therefore split Category I into two mechanistically distinct
sub-clusters: (I.a) structural missing elements --- omitted JSON field,
missing function argument, missing required citation --- which are
absorbed by Category D's constrained-decoding interventions; and (I.b)
semantic missing elements --- incomplete reasoning step, missing edge
case in code, omitted assumption in proof --- which fall under Category
G (process supervision) or Category H (clarification loops). No new
intervention category is required, and two of the highest-volume
clusters (D format + I.a structural missing, jointly \(\approx 25\%\) of
ErrorAtlas) share a single mechanism --- which strengthens rather than
weakens the cluster-orthogonality argument.

Categories B (unit conversion) and C (counting) remain technical gaps.
Unit conversion is naturally folded into Category A: any Python
interpreter handles dimensional analysis, and PAL/PoT/ToRA results
already encompass it. Counting is the smallest residual gap; a focused
literature pass or a one-day inference-only ablation would close it.

\paragraph{Additivity and its limits}\label{additivity-and-its-limits}

Six Sigma Agent results (Six Sigma authors, 2026) demonstrate that
stacking interventions targeting orthogonal failure modes produces
compound reliability gains: in their parallel-consensus framework,
atomic decomposition with five-way voting yields a \(14{,}700\times\)
system reliability improvement over single-pass baselines. AgentSquare
(AgentSquare authors, 2024) similarly shows supra-additive gains when
modules target distinct failure modes. These are encouraging.

There is a caveat. Le (2026) reports that schema-level and prompt-level
instructions interact \emph{non-additively} for some model families:
stacking constrained decoding (Category D) with prompt-level
specification clarifications (Category H) produces less than the sum of
individual gains, sometimes substantially so. The interaction is
concentrated in cases where both interventions share a common prompt
channel. We treat this as evidence not against additivity in general but
for a refinement: interventions that operate on orthogonal processing
layers (decoding constraint vs.~retrieval vs.~training-signal
vs.~inference-time tool call) compose near-additively, while those that
share a channel may interfere. This refinement is itself testable and we
leave it as future work.

\subsubsection{4.3 Claim C: Sublinear length
scaling}\label{claim-c-sublinear-length-scaling}

We now compile direct evidence that LLM reliability decays sublinearly
--- typically logarithmically, sigmoidally, or in a plateau-then-cliff
pattern --- rather than exponentially with output or context length.
This is the headline conclusion of Arbuzov et al.~(2025); we extend it
here with newly available benchmark results and reconcile it with the
steeper-decay reports re-examined in Section 6.

\textbf{Log-linear decay on multi-document reasoning.} Loong (Wang M. et
al., 2024) evaluates 4 task types on 4 context-length sets (10K--50K,
50K--100K, 100K--200K, \textgreater200K). GPT-4o on Chain-of-Reasoning
declines 81.6\% → 62.1\% → 45.1\% → 32.9\% across the four sets. A
log-linear fit accuracy \(\approx 95 - 22 \log_2(L / 10\text{K})\) holds
well; exponential per-token decay would predict near-zero by 200K and is
decisively ruled out.

\textbf{Sigmoid decay on synthetic math.} GSM-\(\infty\) (Zhou et al.,
2025) fits a sigmoid (saturating) shape with \(R^2 > 0.98\) on the
Medium subtask. Two further results from the same paper directly attack
the independent-error model. First, exponentially more inference compute
yields only \emph{linear} AUC gains (\(R^2 > 0.99\) for the linear fit)
--- under independent per-step Bernoulli errors, more trials should give
exponential improvement. Second, DeepSeek-R1 maintains 10\%+ accuracy at
130 reasoning operations; \((1 - \varepsilon)^{130}\) with any
reasonable \(\varepsilon\) would predict near-zero. Both are consistent
with errors concentrated at semantic decision points, not distributed
uniformly.

\textbf{Compositionality gap is bounded, not growing.} Press et
al.~(2023) measure the gap between models' ability to answer two-hop
questions and the product of their one-hop accuracies. Under independent
per-step compounding, the gap should grow with chain length; under
clustered errors, it should be bounded. Press et al.~find a
\(\sim 40\%\) gap that is approximately \emph{constant} across the
entire GPT-3 family from 125M to 175B parameters. This is direct
empirical evidence against per-step independent compounding.

\textbf{Self-consistency gains are too large for independent errors.}
Wang et al.~(2023) report +17.9 pp on GSM8K from majority voting over 40
sampled chains. Under \((1 - \varepsilon)^n\)-style independent errors,
majority voting reduces variance as \(1/\sqrt{m}\) but cannot move the
mean far from the per-step ceiling; the +17.9 pp gain requires correct
answers to function as a narrow attractor in solution space, with errors
dispersed. This is the structural signature of clustered-on-manifold
errors.

\textbf{Bounded \(k\) via context compression.} Pang et al.~(2024)
report that ``Anchor LLMs'' achieve 99\% KV-cache reduction with only
``minor accuracy compromise.'' If long-context reliability depended on
all \(n\) tokens, removing 99\% of them should be catastrophic. The fact
that it is not, plus the 3.5× inference speedup the same paper achieves,
directly implies that the effective decision count \(k\) is much smaller
than \(n\) --- closer to 1\% of context length than to 100\%.
LLMLingua-2 (Jiang et al., 2024) and Recurrent Context Compression
(Huang et al., 2024) report similar bounds.

\textbf{Staircase / plateau-and-cliff patterns.} Think-Prune-Train
(Costello et al., 2025) reports that successive self-training rounds
raise Pass@1 (Gemma-2B: 41.9\% → 57.6\%; LLaMA-70B: 78\% → 91.5\%) while
Pass@20 \emph{plateaus} after the first round. This is the
manifold-transition signature: the model is finding the correct cluster
more reliably (Pass@1 rises) but is not finding new correct solutions
(Pass@20 unchanged). RULER (Hsieh et al., 2024) reports threshold
behaviour: simple retrieval holds to 32K; complex reasoning collapses
earlier. Neither pattern is consistent with smooth
\((1 - \varepsilon)^n\) decay.

\textbf{METR's logistic-in-log-length.} METR (METR, 2025) fits per-model
success as a logistic function of log(human-time):
\(S(t) = \sigma(\beta \cdot (\log t - \log h))\). Logistic-in-log is
mathematically sublinear in length itself. The famous METR exponential
--- capability doubling every seven months --- is an \emph{inter-model}
claim about how the horizon \(h\) grows over generations, not an
intra-model claim about within-model decay shape. We address the common
misreading in Section 6.

\textbf{Counter-evidence.} Several prominent papers report what reads as
rapid or ``exponential'' decay --- Dziri et al.~(2023), Kuratov et
al.~(2024), METR (2025), Wan et al.~(2025). Read carefully, each decays
over a variable other than raw token length: compositional graph size,
number of supporting facts, log-time horizon, capacity threshold.
Section 6 walks through each one.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Practical Implications}\label{practical-implications}

\textbf{Reliability engineering is local patch coverage.} Within a fixed
domain patch, Proposition 1's polylog bound makes reliability a
small-catalogue engineering problem rather than an asymptotic scaling
problem. A team building a reliable LLM system in a measured domain
should initially budget for a library on the order of tens of
interventions, then refine that estimate from the local mode-discovery
curve \(C_{\text{seen},D}(T)\) and the empirical rank-frequency
distribution. Current evidence is consistent with \(\approx 50\) named
interventions covering the head of the per-hard-token failure
distribution in many measured domains --- but this number is a planning
prior, not a universal constant. The same base model in three different
patches (cardiology RAG, legal contract drafting, code review) yields
three different intervention libraries: \(C_D\), \(A_D\), and
\(\sigma_D\) all change with the deployment domain. The engineering goal
is not global manifold mastery; it is identifying the recurring
transition types available in the target patch and provisioning
interventions against the head of that local catalogue.

\textbf{Domain transfer is mild but non-trivial.} The mode-rate constant
\(\sigma\) shifts between domains (math \(\approx 1.2\)--\(1.6\); code
\(\approx 1.0\)--\(1.3\); general \(\approx 1.85\)), and
higher-\(\sigma\) domains carry larger catalogues at any given corpus
size --- so they need proportionally more interventions for matched
coverage. The dependence is linear in \(\sigma\) at fixed coverage: a
domain with twice the mode-discovery rate needs roughly twice the
catalogue size. Across the domains measured here, \(\sigma\) does not
vary by orders of magnitude.

\textbf{Per-hard-token vs.~sequence-level targets matter for SLA
design.} The result in Section 3.4 separates two cases that are often
conflated. Per-hard-token residual error rate --- the right metric for
systems that surface or correct errors continuously --- has a polylog
intervention budget. Sequence-level failure probability --- the right
metric for systems where a single error invalidates the output --- is
strictly stricter and requires nearly full-catalogue coverage as \(k\)
grows. Production deployments should design their reliability SLA to
match the actual cost structure: continuous-correction systems can lean
on the polylog result, while one-shot systems should plan for tighter
coverage.

\textbf{Library design as engineering, not asymptotic theory.} A
practical implication of the cluster-orthogonality finding (Section 4.2)
is that an intervention library can be assembled module-by-module, in
any order, as long as the modules target orthogonal failure modes. The
Le (2026) caveat applies only to interventions sharing a prompt channel;
layer-separated interventions (constrained decoding + retrieval +
process supervision + tool call) compose near-additively. Library
construction is therefore highly parallelisable: different teams can
ship independent interventions and combine them with minimal integration
risk, modulo the prompt-channel caveat.

\textbf{Capability libraries are coarser than error-class libraries.}
Five named error classes in the math domain (arithmetic, units,
counting, list manipulation, date arithmetic) are all addressed by a
single Python interpreter; three code-error classes (SyntaxError,
NameError, most TypeError) collapse under one execution-feedback loop;
format violations and the structural component of ``missing required
element'' --- together more than 20\% of ErrorAtlas --- collapse under
one constrained-decoder deployment. The practical reliability library is
therefore smaller than the count of named error clusters. A team
budgeting for ``\(\approx 50\) interventions per domain'' can expect to
need fewer than 50 capability-axis interventions while still covering
\(\approx 50\) named clusters. The six axes of Section 4.2.0 are likely
closer to the right unit of engineering accounting than the twelve
clusters of Section 4.2.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. Discussion}\label{discussion}

\subsubsection{6.1 By-construction elimination as a boundary
case}\label{by-construction-elimination-as-a-boundary-case}

Six of the 27 GOLD citations in Section 4.2.0 --- DeepSeek-Prover-V2's
Lean proof checker, OpenAI Structured Outputs, ToolDec, DINGO, XGrammar
(×2), and LlmFix's static syntax check --- achieve residual error rate
equal to zero \emph{by mathematical construction}, not statistically.
Constrained decoders set \(P(\text{invalid token}) = 0\) at every
generation step, so no invalid token can be emitted. Proof kernels
reject any output that fails type-checking. Schema validators reject any
output missing a required field. (Acurai's RAG result is excluded from
Pattern A here, despite its 100\% RAGTruth number; it is a
benchmark-conditional empirical observation, not a by-construction
property of citation-grounded retrieval in general --- see Section 4.2.0
Pattern B.)

In this regime, the polylog bound of Proposition 1 ---
\(m \geq |C|^{1 - \varepsilon / e_{\text{hard}}}\) --- is loose. The
cluster covered by a by-construction intervention contributes zero to
the residual error, not a small positive quantity. The formal statement
is that for any cluster \(c \in \{1, \dots, m\}\) covered by a Pattern-A
intervention, \(\Pr(\text{failure} \mid \text{cluster } c) = 0\)
identically, regardless of how many hard-token decisions the model makes
inside that cluster. The Proposition 1 budget formula treats coverage as
a \([0, 1]\) fractional quantity; Pattern A pushes a subset of clusters
to coverage exactly 1.

This is a strengthening result, not a counter-example. The framework's
qualitative claim --- that capability provisioning can erase an error
class --- admits a sharp boundary case where the implication is
mathematical certainty rather than statistical reduction. Crucially,
Pattern A applies precisely to the \emph{structural / verifiable}
classes (format, syntax, schema, invalid proofs), which are also the
head of the heavy-tailed cluster frequency distribution. Capability
provisioning therefore exerts its strongest effect exactly where the
failure-mode catalogue is densest, which compounds the polylog
conclusion of Section 3.4.

\subsubsection{6.2 Re-audit of counter-evidence: relocation, not
dissolution}\label{re-audit-of-counter-evidence-relocation-not-dissolution}

Four prominent papers are routinely cited as evidence that LLM
reliability decays steeply with length, in apparent tension with our
framework: Dziri et al.~(2023), Kuratov et al.~(2024), METR (2025), and
Wan et al.~(2025). To these we add NoCha (Karpinska et al., 2024) for
completeness. A careful re-reading shows that \emph{every one} of these
papers decays over a variable distinct from raw token length \(n\).

Identifying the decay axis is not the same as dissolving the practical
concern. In real tasks where compositional graph size, fact count, or
evidence scope grow with problem length, the framework predicts steep
failure curves; the contribution is to identify which interventions help
(capability provisioning along the actual decay axis) and which do not
(longer context windows, more compute on the \(n\)-axis). The papers
below are not refuted by the framework. Each correctly identifies a
regime in which the framework expects difficulty --- sometimes severe
--- and our analysis below acknowledges those regimes explicitly rather
than relabelling them away.

\textbf{Dziri et al.~(2023), Faith and Fate.} GPT-4 multi-digit
multiplication accuracy drops from 59\% (3-digit) to 4\% (4-digit) in
zero-shot. The authors theorise that ``the probability of incorrect
predictions converges exponentially to \(\approx 1\) for abstract
compositional tasks.'' The decay variable, however, is
\emph{compositional graph size \(N\)}, not raw token length \(n\). The
3×3 multiplication graph has approximately 16 nodes (partial products
plus carries); the 4×4 graph has approximately 25. Multi-digit
multiplication is engineered so that \(k_{\text{hard}} \approx N\) ---
every node in the computation graph is a hard decision. For
natural-language tasks, where the harvest evidence in Section 4.1 gives
\(k_{\text{hard}} \ll n\), the Dziri regime is the \emph{boundary case}
in which our framework reduces to their result. We agree with their
empirical decay; we disagree with the interpretation that it extends to
natural-language \(n\).

A complementary observation: \((1 - \varepsilon)^N = 0.04\) at
\(N = 25\) requires \(\varepsilon \approx 0.12\) per node, but
\((1 - 0.12)^{16} \approx 0.13\), far below the observed 0.59 at
\(N = 16\). Dziri's own data does not fit a clean
\((1 - \varepsilon)^N\) exponential on their own decay axis, suggesting
the actual decay shape is steeper than exponential locally but
compatible with a finite catalogue of failure modes that get exhausted
as \(N\) grows. \textbf{Where the framework concedes}: in adversarial
compositional tasks engineered so that \(k_{\text{hard}} \approx N\),
the framework reduces to Dziri's exponential regime and does not relieve
it; the relocation is informational, not magical.

\textbf{Kuratov et al.~(2024), BABILong.} The abstract reports
``performance declines sharply with increased reasoning complexity'' and
that models ``effectively utilise only 10--20\% of the context.'' Read
in detail, BABILong varies \emph{two} axes: context length \(n\) (from
0K to 10M tokens) and number of supporting facts \(k\) (from QA1 = 1
fact to QA3 = 3 facts, etc.). The sharp decline is in \(k\), not in
\(n\). For QA1 (single-fact), most models ``perform well up to 4,000
tokens'' --- a plateau, not exponential decay. The famous
RAG-flat-across-length result (60\% on single-fact QA \emph{independent
of context length}) is the cleanest possible demonstration that when the
relevant evidence is in window, length does not matter. Recurrent Memory
Transformers maintaining performance to 50M tokens further confirms that
effective \(k\) is determined by architecture, not by raw \(n\).
\textbf{Where the framework concedes}: for multi-hop tasks whose
required fact count grows with task complexity, BABILong's sharp
\(k\)-axis decay is the framework's \emph{predicted} trouble, not a
counter-example to it.

\textbf{METR (2025).} Per-model success fits a logistic curve in
\(\log(\text{human-task-duration})\):
\(S(t) = \sigma(\beta \cdot (\log t - \log h))\). Logistic-in-log-length
is \emph{mathematically sublinear} in length itself. The 80\%-horizon
being 4--6× shorter than the 50\%-horizon is a steep within-model cliff
but is by construction compatible with the polylog result of Section 3.4
--- a cliff at a specific capacity threshold is a manifold-transition
signature, exactly the structure of Arbuzov et al.~(2025, §3.2). The
famous exponential --- capability doubling every seven months --- is an
\emph{inter-model} claim about how the horizon \(h\) moves across
generations, orthogonal to within-model decay shape. When a model
improves, its \(h\) shifts right; the curve does not steepen.
\textbf{Where the framework concedes}: the within-model logistic-in-log
cliff is steep on a practical scale; calling it ``sublinear in length''
is technically correct but engineering-useful only for systems that can
be designed to operate well below the 50\% horizon.

\textbf{Wan et al.~(2025), Fano-style upper bound.} This paper theorises
super-linear information-demand growth and identifies an ``accuracy
cliff'' at capacity overflow: ``when the task's information demand
surpasses the model's output capacity, performance does not degrade
gracefully but instead collapses sharply.'' The behavioural prediction
is a \emph{threshold}, not smooth \((1 - \varepsilon)^n\). A cliff at a
specific capacity threshold is exactly the manifold-transition behaviour
our Section 3 framework predicts at the boundary between covered and
uncovered modes; Wan et al.'s mechanism (capacity overflow) is one
specific cause of the boundary, complementary to ours. \textbf{Where the
framework concedes}: Wan documents a mechanism by which \(|C|\)
effectively exceeds the model's coverage in a single forward pass;
Pattern A interventions (constrained decoding, formal verification) are
the kind of response the framework expects but does not automatically
supply.

\textbf{Karpinska et al.~(2024), NoCha.} GPT-4o achieves 55.8\% pair
accuracy across 1,001 minimally-different true/false claim pairs about
long fictional books (mean length 127K tokens). The paper reports
performance by \emph{evidence scope}, not by context length: 59.8\% on
sentence-level retrieval, 47.6\% on passage-level, 41.6\% on global
reasoning. No per-context-length curve is reported. The decay axis is
the number of evidence pieces that must be integrated --- i.e., a
\(k\)-axis quantity, not an \(n\)-axis one. NoCha is therefore not
counter-evidence to a sublinear-in-\(n\) claim. \textbf{Where the
framework concedes}: evidence-scope decay is a \(k\)-axis observation
that the framework predicts cannot be addressed by scaling raw context
length; it requires retrieval or process supervision along the actual
decay axis.

\textbf{Unifying observation.} Every counter-paper decays over a
variable that is not \(n\): compositional graph size, fact count,
log-time horizon, capacity threshold, or evidence scope. The apparent
rapid decay is, in each case, in a quantity that our framework already
concentrates the action in (\(k_{\text{hard}}\) and \(|C|\)), not in raw
sequence length. This is a relocation, not a dissolution. These papers
are not counter-evidence to the sublinear-in-\(n\) thesis, but they
identify regimes where the framework predicts steep
\(k_{\text{hard}}\)-axis decay in practice. Identifying the correct
intervention axis is the contribution of this re-audit; promising that
those regimes are easy is not. Where \(k_{\text{hard}}\) grows with task
length --- adversarial compositional structure, multi-hop fact chains,
long horizons that force more decisions --- reliability remains hard;
the framework's value is in directing intervention toward capability
provisioning along the actual decay axis rather than toward
context-window or compute-budget expansion that does not help.

\subsubsection{6.3 Limitations, residual budget, and
falsifiability}\label{limitations-residual-budget-and-falsifiability}

\textbf{Mathematical limitations.} Three are worth flagging. First,
Postulate 1 is empirical, not derived. While it is consistent with
ErrorAtlas, HumanEval, and MWPES at corpus scale, a domain with
genuinely Heaps-power-law mode discovery (canonical \(b \approx 0.5\))
would invalidate the doubly-logarithmic special case of Proposition 1,
although the qualitative polylog result of Section 3.5 would still hold.
Second, the coverage form in Section 3.3 is an empirical best-fit choice
among several plausible candidates; readers should not interpret the
close numerical match against two anchor points as theoretical
confirmation. Third, the cluster-orthogonality assumption that
underwrites Proposition 1's composition step is only approximate; the Le
(2026) caveat on prompt-channel interference is real and would tighten
the bound in shared-channel settings.

\textbf{Domain narrowness.} Postulate 1's empirical anchor rests on
three taxonomies (ErrorAtlas, HumanEval categorisation, MWPES-300K) all
published in 2025--2026 and covering general, code, and math domains.
The framework is untested on agentic workflows, long-running scientific
reasoning, multi-turn tool use over millions of tokens, and production
codebases. Those are precisely the regimes where \(|C|\) may grow faster
than logarithmically --- and where the practical reliability question
matters most. The polylog conclusion of Section 3.4 should be read as a
domain-conditional claim until subsample-discovery measurements (Section
3.6 lists the falsifiability test) are available in those regimes.

\textbf{Single-point calibration of \(\sigma\).} The estimate
\(\sigma \approx 1.85\) is a calibration against the ErrorAtlas category
count at one corpus scale, not a fit to a discovery curve. A
subsample-vs-distinct-modes plot has not been published for ErrorAtlas
(or for any other LLM error taxonomy at the time of writing); Section
3.6 lists this as the explicit empirical test that would either tighten
the postulate or move the analysis to the Heaps variant of Section 3.5.

\textbf{Identifiability of \(\beta\).} \(\beta\) is a latent
stratification parameter; we report no empirical range because
failure-mode atlases measure
\(\Pr(\text{category} \mid \text{error occurred})\), which does not
identify \(\Pr(\text{hard} \mid \text{key-token})\). Future work
measuring hard-token opportunities directly --- e.g., by ablation of
model confidence at predicted decision boundaries, or by comparing
confidence distributions over sampled key-token positions --- would
close this gap.

\textbf{Taxonomy granularity and the L2--L3 gap.} The framework operates
on L2 (taxonomy categories) as a proxy for L3 (latent modes). \(|C|\)
depends on the taxonomer's resolution choices: splitting ``reasoning
error'' into 200 subtypes inflates \(|C|\) without changing the
underlying mode catalogue. We have no independent measurement of how
faithful current taxonomies are as L3 proxies, and we treat this as a
residual epistemic risk rather than a solved problem.

\textbf{Goalpost-relocation caveat.} By concentrating the practical
action in \(k_{\text{hard}}\) and \(|C|\) rather than raw \(n\), the
framework \emph{relocates} the difficulty of long-context reliability
rather than resolving it. Domains where \(k_{\text{hard}}\) grows with
task length --- adversarial compositional structure, multi-hop chains,
long agent horizons --- remain hard. The framework's claim is to
identify the axis of intervention (capability provisioning along the
actual decay variable), not to dissolve the underlying engineering
problem. Practical reliability work in those regimes still requires the
hard interventions; the framework's value is in saying which
interventions are on-axis and which are off.

\textbf{Patch-shift between deployments.} When a model moves from one
application domain to another --- say, from general-purpose chat to
clinical decision support --- the four domain-indexed quantities
\(A_D\), \(\sigma_D\), \(\beta_D\), and \(|C_D|\) all change. A library
calibrated against one patch will under-cover the next patch's mode
catalogue. The framework predicts this transfer failure structurally
(heavy-tailed and bounded per patch) but says nothing about the
\emph{content} of \(C_{D'}\) in advance. The operationally useful
response is to re-measure inside the new patch and re-fit the head of
the local distribution. The cleanest follow-up evidence would be
per-domain mode-discovery measurement --- subsamples of size \(t\)
versus distinct discovered modes, the falsifiability test of §3.6 ---
but no such curve has been published at the time of writing. The
framework's prediction is that such curves should rise log- or
Heaps-style inside the patch and asymptote to a domain-specific
\(|C_D|\).

\textbf{The irreducible-semantic residual.} Of the 17 ErrorAtlas
categories, 13 are directly addressed by one of the six capability axes
of Section 4.2.0 --- under Patterns A, B, or C. The remaining four ---
inappropriate refusal (partially covered by preference optimisation),
specification misinterpretation (Category H, addressed only weakly by
clarification loops), reasoning bottlenecks at the level of problem
decomposition, and a small residual of ``user wanted something
different'' semantic failures --- admit no clean capability-provisioning
fix in the current literature. These are \emph{irreducible-semantic
residuals}: classes where the failure is in choosing what to do, not in
executing it. Proposition 1's prediction of \(O(\log)\) rather than
\(O(0)\) residual reliability reflects exactly this irreducible core.
The framework does not claim 100\% coverage of all failures; it claims
polylog-bounded \emph{capability-eliminable} failures, with the named
semantic residual as the floor.

\textbf{A falsifiability test the framework passes.} DebugBench (Tian et
al., 2024) explicitly catalogues which error classes execution-feedback
can and cannot repair. It works well for syntax and reference errors ---
quoted: \emph{``SyntaxError can be fixed: 99.74\% on HumanEval, 100.00\%
on MBPP.''} It does not work for logic errors --- quoted: \emph{``the
feedback information is unhelpful for logic errors\ldots{} may even
cause disruptions.''} This is the framework's predicted failure mode
(execution feedback is a Pattern-A/B capability for structural-class
errors and a Pattern-C-at-best for semantic-class errors) and a
falsifiability test it passes. If capability provisioning bled into the
semantic residual --- if execution feedback claimed to fix logic errors
--- the framework's selectivity claim would be wrong. The empirical
result is that capability provisioning correctly identifies both what it
can erase and what it cannot.

\textbf{Net assessment.} The framework correctly characterises both the
success regime and the failure regime; the residuals are honest
limitations, not silent failures. A future extension addressing the
irreducible-semantic residual --- through richer process supervision,
ensemble-based verification, or higher-order capability layers --- would
tighten the bound but is not required for the polylog conclusion to hold
over the capability-eliminable fraction.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{7. Conclusion}\label{conclusion}

LLM reliability is often framed as an asymptotic scaling problem: as
outputs get longer, do errors compound to inevitable failure? In our
previous work (Arbuzov et al., 2025) we argued that this framing rests
on a false uniformity assumption --- errors concentrate at
\(\sim 5{-}10\%\) of tokens, not all of them. This paper takes the next
step: within that sparse set, errors are not only concentrated but also
\emph{repetitive}. They cluster into a finite catalogue of recurring
failure modes whose size, under the empirical postulate of Section 3.2,
grows logarithmically --- and under the more conservative Heaps
alternative, as a small power --- of the number of observed failures.

The conditional consequence (Proposition 1) is that the intervention
budget required to bound per-hard-token residual error scales
polylogarithmically in sequence length within a fixed domain patch, and
becomes a domain-constant once the patch ceiling \(|C_D|\) is reached.
The optimistic special case (Postulate 1, \(k \sim \log n\)) gives a
doubly-logarithmic rate; the conservative Heaps variant of Section 3.5
gives \((\log n)^b\) with small \(b\); the symbolic-rates form of
Section 3.6 shows the polylog conclusion is stable across this family.
Available evidence is consistent with libraries on the order of tens of
interventions covering the head of the per-hard-token failure
distribution in many fixed domains. The exact number is patch-indexed
and should be estimated from local discovery and rank-coverage curves,
not assumed from cross-domain priors. Sequence-level reliability targets
are strictly tighter and approach full-catalogue coverage as \(k\)
grows; practitioners should pick the regime that matches their cost
structure rather than reading the per-token result as the production
SLA.

This reframes the long-context reliability question from ``can we bound
the growth of \(n\)-token error?'' to ``have we catalogued enough
failure modes?'' The latter question is finite and addressable. Whether
it generalises to agentic, scientific, and long-horizon regimes where
the published taxonomies do not yet reach is the open empirical question
this paper invites. In the domains where taxonomies have been measured
--- general, code, math --- the architecture of LLM errors is compact
and the engineering problem is small. Whether the same holds elsewhere
is a falsifiable empirical test, and we have named it.

Several extensions stand out. Direct empirical measurement of the
mode-rate constant \(\sigma\) on new domains --- agentic, scientific,
code-in-production --- would test the cross-domain generality of
Postulate 1 and surface outlier domains where the catalogue may grow
faster than logarithmically. A sequence-level analogue of Proposition 1,
replacing the per-hard-token bound with a probabilistic bound on
sequence failure, would close the gap between Section 3.4's two regimes
and give one-shot systems a cleaner SLA target. The deepest question is
structural: why does mode discovery look logarithmic in the first place?
Integration with the attention-mechanism bounds and stratified-manifold
geometry flagged as future work in Arbuzov et al.~(2025, §6) would
replace our empirical postulate with a derived result --- closing the
loop between the two papers.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{References}\label{references}

Acurai authors. (2024). Acurai: Query reformatting for
hallucination-free retrieval-augmented generation. \emph{arXiv preprint
arXiv:2412.05223}.

AgentSquare authors. (2024). AgentSquare: Modular composition for LLM
agents. \emph{arXiv preprint arXiv:2410.06153}.

Almanac authors. (2024). Almanac: Retrieval-augmented large language
models for clinical question answering. \emph{NEJM AI}.

Arbuzov, M. L., Shvets, A. A., \& Bei, S. (2025). Beyond exponential
decay: Rethinking error accumulation in large language models.
\emph{arXiv preprint arXiv:2505.24187}.

Asai, A., et al.~(2024). Self-RAG: Learning to retrieve, generate, and
critique through self-reflection. \emph{International Conference on
Learning Representations (ICLR) 2024}. arXiv:2310.11511.

Ashury-Tahan, Y., et al.~(2026). ErrorAtlas: A large-scale empirical
atlas of LLM failure patterns across models and benchmarks. \emph{arXiv
preprint arXiv:2601.15812}.

Chen, W., Ma, X., Wang, X., \& Cohen, W. W. (2022). Program of thoughts
prompting: Disentangling computation from reasoning for numerical
reasoning tasks. \emph{arXiv preprint arXiv:2211.12588}.

Cheng, K., et al.~(2024). SeeClick: Harnessing GUI grounding for
advanced visual GUI agents. \emph{Proceedings of ACL 2024}.
arXiv:2401.10935.

Costello, C., Wells, C., Grefenstette, E., \& Glaese, A. (2025). Think,
prune, train, improve: Scaling reasoning without scaling models.
\emph{arXiv preprint arXiv:2504.18116}.

Dai, X., et al.~(2025). EDIT: Identifying key reasoning steps in
chain-of-thought. \emph{Proceedings of ACL 2025}. arXiv:2405.19737.

Dantart, J. (2026). The rigorous archivist: Source-restricted retrieval
for fabrication-free legal citation. \emph{arXiv preprint
arXiv:2601.15476}.

DeepSeek-Prover team. (2025). DeepSeek-Prover-V2: Advancing formal
mathematical reasoning via large-scale reinforcement learning with
verifier feedback. \emph{arXiv preprint arXiv:2504.21801}.

Dong, Y., et al.~(2024). XGrammar: Efficient grammar-constrained
decoding for large language models. \emph{arXiv preprint
arXiv:2411.15100}.

Dong, Y., et al.~(2026). XGrammar-2: Scaling schema-constrained tool
calls. \emph{arXiv preprint arXiv:2601.04426}.

Dziri, N., Lu, X., Sclar, M., Li, X. L., Jiang, L., Lin, B. Y., West,
P., Bhagavatula, C., Le Bras, R., Hwang, J. D., et al.~(2023). Faith and
fate: Limits of transformers on compositionality. \emph{Advances in
Neural Information Processing Systems 36}. arXiv:2305.18654.

Fang, L., Wang, Y., Liu, Z., Zhang, C., Jegelka, S., Gao, J., Ding, B.,
\& Wang, Y. (2024). What is wrong with perplexity for long-context
language modeling? \emph{International Conference on Learning
Representations (ICLR) 2025}. arXiv:2410.23771.

Gao, L., Madaan, A., Zhou, S., Alon, U., Liu, P., Yang, Y., Callan, J.,
\& Neubig, G. (2022). PAL: Program-aided language models.
\emph{Proceedings of ICML 2023}. arXiv:2211.10435.

Gao, T., et al.~(2023). ALCE: Enabling large language models to generate
text with citations. \emph{Proceedings of EMNLP 2023}. arXiv:2305.14627.

Geng, S., et al.~(2025). JSONSchemaBench: Evaluating constrained
decoding. \emph{arXiv preprint arXiv:2501.10868}.

Goodell, A., et al.~(2025). Task-specific tools for clinical
calculations: A large language model evaluation. \emph{Nature npj
Digital Medicine}.

Gou, B., et al.~(2025). UGround: Navigating the digital world as humans
do. \emph{International Conference on Learning Representations (ICLR)
2025}. arXiv:2410.05243.

Gou, Z., et al.~(2023). ToRA: A tool-integrated reasoning agent for
mathematical problem solving. \emph{arXiv preprint arXiv:2309.17452}.

Gou, Z., et al.~(2024). CRITIC: Large language models can self-correct
with tool-interactive critiquing. \emph{International Conference on
Learning Representations (ICLR) 2024}. arXiv:2305.11738.

Hsieh, C.-P., Sun, S., Kriman, S., Acharya, S., Rekesh, D., Jia, F.,
Zhang, Y., \& Ginsburg, B. (2024). RULER: What's the real context size
of your long-context language models? \emph{arXiv preprint
arXiv:2404.06654}.

Huang, C., Zhu, G., Wang, X., Luo, Y., Ge, G., Chen, H., Yi, D., \&
Wang, J. (2024). Recurrent context compression: Efficiently expanding
the context window of LLM. \emph{arXiv preprint arXiv:2406.06110}.

Huang, D., et al.~(2023). AgentCoder: Multi-agent-based code generation
with iterative testing and optimisation. \emph{arXiv preprint
arXiv:2312.13010}.

Jiang, H., Wu, Q., Luo, X., Li, D., Lin, C.-Y., Yang, Y., \& Qiu, L.
(2024). LongLLMLingua: Accelerating and enhancing LLMs in long context
scenarios via prompt compression. \emph{arXiv preprint
arXiv:2310.06839}.

Karaman, O. M., et al.~(2024). POROver: Improving safety and reducing
over-refusal in large language models with over-refusal-aware preference
optimization. \emph{arXiv preprint arXiv:2410.12999}.

Kandpal, N., Deng, H., Roberts, A., Wallace, E., \& Raffel, C. (2023).
Large language models struggle to learn long-tail knowledge.
\emph{Proceedings of ACL 2023}. arXiv:2211.08411.

Karpinska, M., et al.~(2024). One thousand and one pairs: A ``novel''
challenge for long-context language models. \emph{Proceedings of EMNLP
2024}. arXiv:2406.16264.

Kuratov, Y., Bulatov, A., Anokhin, P., Rodkin, I., Sorokin, D., Sorokin,
A., \& Burtsev, M. (2024). BABILong: Testing the limits of LLMs with
long context reasoning-in-a-haystack. \emph{NeurIPS 2024 Datasets and
Benchmarks Track}. arXiv:2406.10149.

Le, T. (2026). Schema-level and prompt-level instructions interact
non-additively in structured generation. \emph{arXiv preprint}.

LeCun, Y. (2023). Why autoregressive language models are exponentially
diverging and doomed. \emph{Public commentary}.

Li, X., \& Sarwate, A. D. (2025). Unraveling the localized latents:
Learning stratified manifold structures in LLM embedding space with
sparse mixture-of-experts. \emph{arXiv preprint arXiv:2502.13577}.

Li, Y., et al.~(2022). Competition-level code generation with AlphaCode.
\emph{Science}, 378(6624), 1092--1097.

Lightman, H., et al.~(2023). Let's verify step by step. \emph{arXiv
preprint arXiv:2305.20050}.

Liu, F., et al.~(2023). DePlot: One-shot visual language reasoning by
plot-to-table translation. \emph{Findings of ACL 2023}.
arXiv:2212.10505.

Lu, Y., et al.~(2024). OmniParser for pure vision based GUI agent.
\emph{arXiv preprint arXiv:2408.00203}.

Mallen, A., et al.~(2023). When not to trust language models:
Investigating effectiveness of parametric and non-parametric memories
(PopQA). \emph{Proceedings of ACL 2023}. arXiv:2212.10511.

Manning, C. D., Raghavan, P., \& Schütze, H. (2008). \emph{Introduction
to information retrieval}. Cambridge University Press. (Section on
Heaps' law:
\url{https://nlp.stanford.edu/IR-book/html/htmledition/heaps-law-estimating-the-number-of-terms-1.html}.)

MAST authors. (2025). MAST: Failure-mode analysis of multi-agent LLM
systems. \emph{arXiv preprint arXiv:2503.13657}.

METR. (2025). Measuring AI ability to complete long software tasks.
\emph{arXiv preprint arXiv:2503.14499}.

Niwa, S., \& Iso, H. (2024). AmbigNLG: Addressing task ambiguity in
open-domain NLG. \emph{Proceedings of EMNLP 2024}. arXiv:2402.17717.

OpenAI. (2024). Introducing structured outputs in the API. \emph{OpenAI
Engineering Blog}.
\url{https://openai.com/index/introducing-structured-outputs-in-the-api/}.

OpenAI. (2025). GPT-5 system card. \emph{OpenAI}.
\url{https://openai.com/index/gpt-5-system-card/}.

Pang, J., Ye, F., Wong, D. F., He, X., Chen, W., \& Wang, L. (2024).
Anchor-based large language models. \emph{Findings of ACL 2024}.
arXiv:2402.07616.

Park, K., Choe, Y. J., \& Veitch, V. (2024). The linear representation
hypothesis and the geometry of large language models. \emph{Proceedings
of ICML 2024}. arXiv:2311.03658.

Press, O., Zhang, M., Min, S., Schmidt, L., Smith, N. A., \& Lewis, M.
(2023). Measuring and narrowing the compositionality gap in language
models. \emph{Findings of EMNLP 2023}. arXiv:2210.03350.

RFM authors. (2025). RFMDataset: Recurring failure modes across
open-source LLMs. \emph{arXiv preprint arXiv:2506.17114}.

Schaeffer, R., et al.~(2025). Why are large language models' scaling
laws power laws? \emph{International Conference on Machine Learning
(ICML) 2025}. arXiv:2502.17578.

Schick, T., et al.~(2023). Toolformer: Language models can teach
themselves to use tools. \emph{Advances in Neural Information Processing
Systems 36}. arXiv:2302.04761.

Shi, F., et al.~(2024). MGDebugger: From syntax to semantics ---
multi-granularity LLM-driven code debugging. \emph{arXiv preprint
arXiv:2410.01215}.

Shinn, N., Cassano, F., Berman, E., Gopinath, A., Narasimhan, K., \&
Yao, S. (2023). Reflexion: Language agents with verbal reinforcement
learning. \emph{Advances in Neural Information Processing Systems 36}.
arXiv:2303.11366.

Six Sigma authors. (2026). Six Sigma agent framework: Compound
reliability through orthogonal interventions. \emph{arXiv preprint}.

SLOT authors. (2025). SLOT: Schema-level output transformation for
reliable structured generation. \emph{arXiv preprint arXiv:2505.04016}.

Sun, J., et al.~(2025). MWPES-300K: A large-scale math word problem
error set with per-category failure distributions. \emph{Findings of ACL
2025}. arXiv:2501.15581.

Suresh, S., et al.~(2025). DINGO: Constrained decoding for diffusion
language models. \emph{arXiv preprint arXiv:2505.23061}.

Suri, K., et al.~(2025). SAGE-Agent: Structured uncertainty in agentic
AI. \emph{arXiv preprint arXiv:2511.08798}.

TauBench authors. (2024). \(\tau\)-bench: A benchmark for tool-augmented
LLM agents. \emph{arXiv preprint arXiv:2406.12045}.

Tian, R., et al.~(2024). DebugBench: Evaluating debugging capability of
large language models. \emph{arXiv preprint arXiv:2401.04621}.

Wada, S., et al.~(2025). Domain retrieval-augmented generation
eliminates radiology report hallucinations. \emph{Nature npj Digital
Medicine}.

Wan, K., et al.~(2025). A Fano-style accuracy upper bound for LLM
single-pass reasoning in multi-hop QA. \emph{arXiv preprint
arXiv:2509.21199}.

Wang, M., et al.~(2024). Leave no document behind: Benchmarking
long-context LLMs with extended multi-doc QA. \emph{Proceedings of EMNLP
2024}. arXiv:2406.17419.

Wang, P., et al.~(2023). Math-Shepherd: Verify and reinforce LLMs
step-by-step without human annotations. \emph{arXiv preprint
arXiv:2312.08935}.

Wang, X., Wei, J., Schuurmans, D., Le, Q., Chi, E., Narang, S.,
Chowdhery, A., \& Zhou, D. (2023). Self-consistency improves chain of
thought reasoning in language models. \emph{International Conference on
Learning Representations (ICLR) 2023}. arXiv:2203.11171.

Wang, Y., et al.~(2024). MMLU-Pro: A more robust and challenging
multi-task language understanding benchmark. \emph{arXiv preprint
arXiv:2406.01574}.

Wang, Z., et al.~(2025). From scores to steps: Diagnosing and improving
LLM performance on medical calculation tasks (MedRaC). \emph{Proceedings
of EMNLP 2025}. arXiv:2509.16584.

Weakest Link authors. (2026). The weakest link law: Multi-hop QA
performance collapses to the least-visible-evidence bucket. \emph{arXiv
preprint arXiv:2601.12499}.

Xie, T., et al.~(2025). Jedi: A grounding-dataset-driven approach to
OSWorld and GUI agent tasks. \emph{NeurIPS 2025}. arXiv:2509.13644.

Zhang, J., et al.~(2024). Categorizing code-generation errors of large
language models on HumanEval (LlmFix). \emph{arXiv preprint
arXiv:2409.00676}.

Zhang, K., et al.~(2023). ToolDec: Syntax error-free and generalisable
tool use via finite-state-machine constrained decoding. \emph{arXiv
preprint arXiv:2310.07075}.

Zhou, Y., Liu, H., Chen, Z., Tian, Y., \& Chen, B. (2025). GSM-Infinite:
How do your LLMs behave over infinitely increasing context length and
reasoning complexity? \emph{arXiv preprint arXiv:2502.05252}.

Zhu, H., et al.~(2025). Lost-in-retrieval: Decomposition failures in
multi-hop retrieval-augmented generation. \emph{arXiv preprint
arXiv:2502.14245}.

\end{document}
